### 2016-11-26 abbReturns_Script.Rmd
filled out data prep

---
title: "Should I AirBNB My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "1 November 2016"
output:
  html_notebook: default
  pdf_document: default
---

&nbsp;

## Introduction

&nbsp;

This document is intended the literate programming of our analysis for the 'Should I AirBNB My Property?' research project.  Below we document the complete data provenance, taking our data from the raw source data through to the final results. Data prepartion, modeling and visualization will be done in R (version 3.3.1). As a very brief overview, we will be comparing the financial returns of short-term (AirBNB-type) and long-term (traditional leases) leasing of residential properties in Melbourne, Australia.  We perform this comparison under a hypothetical condition where-in an investor purchases a property on Septemeber 1, 2015 and is looking to determine whether too lease out the property on a traditional long-term lease (12 months) or as an AirBNB property.  We examine the returns for the following 12 months.  Sensitivity to time frames, start dates, locations, property types, etc. are considered in this analysis as well.

Data on AirBNB (short-term) ltrals has been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company.

Throughout this document we will refer to the short-term data as **'short-term'** or **'AirBNB'** while the long-term data will be referred to as **'long-term'**, **'ltral'** or **'lease'**.  Short-term nightly fees will be referenced as **'rates'** and long-term as **'ltr'**.  We will use **'prices'** to collectively refer to **rates** and **ltrs**. Note that within the Melbourne market ltrals are quoted in weekly ltrs (but are charged to tenants at a monthly ltr equivalent to the 1/12 of 52 weeks of ltr).

&nbsp;

### Preliminary Commands

&nbsp;

We begin with a number of preliminary commands that include setting the global options, loading necessary packages (libraries), setting the path to the raw data sources and reading in custom functions used later in the analysis.

```{r prelim_cmds, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

 ## Set global options

  options(xtable.comment = FALSE)

 ## Load libraries

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)

 ## Set path location

  # Get computer names
  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name
  if(comp.name == '7020D-121777-W' | 
      comp.name == 'DESKTOP-1D7JO4J'){

    data.path <- 'c:/dropbox/research/airBNB/data/'

  } else {

    data.path <- 'gideon path'

  }

 ## Load sources (custom function files)

  source("c:/code/dataviztools/ggplottools.R")
  source("c:/code/datamgmttools/dataMungetools.R")
  source("c:/code/research/AirBNBMelbourne/analysis_Functions.R")
  source("c:/code/research/AirBNBMelbourne/dataPrep_Functions.R")

 ## Set a global parameter

  exch.rate <- 1.32

```

```{r set_colors, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

 ## Set basic color to use in palettes

  abb.colors.names=c('rausch', 'hackberry', 'kazan', 'babu', 'lima', 'beach',
                     'ebisu', 'tirol', 'foggy', 'hof')
  abb.col <- c(rgb(250,88,99, maxColorValue=255), rgb(139,14,82, maxColorValue=255),
               rgb(0,117,140, maxColorValue=255), rgb(4,211,191, maxColorValue=255), 
               rgb(77,226,110, maxColorValue=255), rgb(252,179,14, maxColorValue=255),
               rgb(249,177,139, maxColorValue=255), rgb(191,166,100, maxColorValue=255),
               rgb(156,161,155, maxColorValue=255), rgb(86,94,97, maxColorValue=255))

  sm.col <- abb.col[c(1, 3, 6, 5, 2)]

```

&nbsp;

### Data Pre-processing

&nbsp;

Both the AirBNB and the ltral data require some measure of data pre-processing and due to the run times of these function they are done outside of this document.  For the AirBNB data, the pre-processing is primarily concerned with imputing missing daily observations for a small number of properties in the dataset.  Full information on this process can be found in the **airbnbDataClean.Rmd** and **dataCleanFunctions.R** files.  Do note that most of the imputed data is ignored by our choice of study time period later on.

The long term data received from the data provider is both wide and long and very inefficiently stored.  To aid in computation time, the pre-processing involves the trimming of fields, labeling of observation types and reconfiguration of data and can be found in the **fixAPMData.R** script.

&nbsp;

## Loading data

&nbsp;

Next we load in the raw data.  The short-term data is found in two separate files, one providing information on the property itself data and one providing a daily record of bookings.

```{r load_abb_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Property Information

  abb.data <- read.csv(paste0(data.path, '/prepared/propdata.csv'), header=T)

  # Daily ltral information

  daily.data  <- read.csv(paste0(data.path, '/prepared/dailydata.csv'), header=T)

```

We then load the long term ltral data.  This includes property level data (including the final lease rate) and listing-level data containing information on all changes to listing prices and dates, etc.

```{r load_ltr_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Property level long term ltral information

  ltr.data <- read.csv(paste0(data.path, '/prepared/ltpropdata.csv'), header=T)

 ## Long term ltral listing information

  listing.data <- read.csv(paste0(data.path, '/prepared/ltlistdata.csv'), header=T)

```

And, finally, we load two GIS shapefiles that denote suburb and Statistical Area 1 (sa1) boundaries in the Melbourne area. We also load a shapefile that contains the location of the beach areas along Port Phillip Bay.^[Only those east of the mouth of Yarra River.]

```{r load_sp_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## SA1 shapefile

  sa1s.shp <- readShapePoly(paste0(data.path, 'geographic/sa1s.shp'),
                            proj4string=CRS("+init=epsg:4283"),
                            delete_null_obj=TRUE)

  ## Suburb shapefile

  suburbs.shp <- readShapePoly(paste0(data.path, 'geographic/melbSuburbs.shp'),
                               proj4string=CRS("+init=epsg:4283"),
                               delete_null_obj=TRUE)

  ## Beach file

  beach.shp <- readShapePoly(paste0(data.path, 'geographic/portPhillipBeach.shp'),
                             proj4string=CRS("+init=epsg:4283"),
                             delete_null_obj=TRUE)

```

&nbsp;

## Data Cleaning and Preparation

&nbsp;

In this section we remove, convert and/or transform a number of data fields.  We also remove observations if they are missing critical data or fall outside of the scope of our research question.

&nbsp;

### Trim Fields

&nbsp;

We start by converting all field names to lower case to simplify future coding.  This is important as R is case sensitive.

```{r lower_names, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Short term data

  names(abb.data) <- tolower(names(abb.data))
  names(daily.data) <- tolower(names(daily.data))

 ## Long term data

  names(ltr.data) <- tolower(names(ltr.data))
  names(listing.data) <- tolower(names(listing.data))

 ## Geographic data

  names(sa1s.shp@data) <- tolower(names(sa1s.shp@data))
  names(suburbs.shp@data) <- tolower(names(suburbs.shp@data))

```

The data contain many fields that are either not fully observed and/or are not relevant to our research question.  Additionally, the field naming conventions differ between the two datasets.  We first remove unnecessary or unobserved fields and then employ a systematic field naming functions across all datasets.

We start by trimming the property-level short term data.  No fields need to be removed from the daily data.

```{r trim_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Property Data

  abb.keep.flds <- c('property.id', 'host.id', 'listing.title', 'property.type',
                     'listing.type', 'created.date', 'latitude', 'longitude',
                     'last.scraped.date', 'average.daily.rate', 'overall.rating',
                     'max.guests', 'cancellation.policy', 'cleaning.fee',
                     'published.nightly.rate', 'published.weekly.rate', 
                     'published.monthly.rate', 'minimum.stay',
                     'bedrooms', 'bathrooms', 'min.date', 'max.date', 'total.days',
                     'blocked.days', 'available.days', 'created.year')

  abb.data <- abb.data[ , abb.keep.flds]

```

We then limit the fields of the long-term property-specific data

```{r trim_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Long term property Data

  ltr.keep.flds <- c('id.key', 'geographicalid', 'eventid', 'addressid', 'activityid',
                      'flatnumber', 'streetnumber', 'streetname', 'streettype', 
                      'suburb', 'postcode', 'property_latitude',
                      'property_longitude', 'street_centroid_latitude',
                      'street_centroid_longitude', 'eventdate', 'eventprice',
                      'propertytype', 'areasize', 'bedrooms', 'baths', 'parking',
                      'hasstudy', 'hascourtyard', 'hasbalcony', 'hasairconditioning',
                      'hasgarage', 'dom')

  ltr.data <- ltr.data[ , ltr.keep.flds]

```

&nbsp;

### Standardize Field Names

&nbsp;

Next we standardize the field names.  Here we follow the short-term data convention of placing a period between words in the name.  We also shorten a number of the fields to make future coding easier.  No changes are made to the monthly or daily data.

For the short term data, it already uses the *word-.-word* convention.  Here our renaming is done to shorten certain field names.

```{r rename_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(abb.data) <- c('property.id', 'host.id', 'listing.title', 'property.type',
                       'listing.type', 'created.date', 'latitude', 'longitude',
                       'last.scraped.date', 'average.daily.rate', 'overall.rating',
                       'max.guests', 'cancel.policy', 'cleaning.fee',
                       'nightly.rate', 'weekly.rate', 'monthly.rate', 'min.stay',
                       'bedrooms', 'baths', 'min.date', 'max.date', 'total.days',
                       'blocked.days', 'available.days', 'created.year')

```

Periods are inserted between relevant words for the ltral data.

```{r rename_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(ltr.data) <- c('id.key', 'geographical.id', 
                        'event.id', 'address.id', 'activity.id',
                        'flat.number', 'street.number', 'street.name', 'street.type',
                        'suburb', 'postcode', 'latitude', 'longitude',
                        'street.latitude', 'street.longitude', 'event.date',
                        'event.price', 'property.type', 'area.size', 'bedrooms',
                        'baths', 'parking', 'has.study', 'has.courtyard',
                        'has.balcony', 'has.aircond', 'has.garage', 'dom')

  names(listing.data) <- c('id.key', 'event.id', 'activity.id', 'address.id', 
                           'of.record',
                           'event.date', 'event.price', 'first.adv.date', 
                           'first.adv.price', 'last.adv.date', 'last.adv.price',
                           'event.type')

```

&nbsp;

### Field Conversions

&nbsp;

Below we convert and coalate a number of the temporal and spatial variables.

&nbsp;

#### Convert Time Variables

We convert time variables across all dataset to an R date format.  For the AirBNB data this is a relatively straighforword process as the data are already standardized into a consistant date representation. We also create a year variable.

```{r fix_st_time, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Property Data Fields

  abb.data$created.date <- as.Date(abb.data$created.date)
  abb.data$last.scraped.date <- as.Date(abb.data$last.scraped.date)
  abb.data$min.date <- as.Date(abb.data$min.date)
  abb.data$max.date <- as.Date(abb.data$max.date)

  ## Daily Data

  daily.data$date <- as.Date(daily.data$date)

```

The dates in the long term data are not standardized.  To fix this issue we create a small function that will remove the unnecessary digits and spaces and then convert to a standard date form.  We then apply this formula to the three date fields in the long term data.  We also create a year field for the event date.

```{r fix_lt_time, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Build custom function

  fixLTDates <- function(x){
    temp.date <- as.character(x)
    temp.date <- str_replace_all(temp.date, ' 0:00', '')
    temp.date <- str_replace_all(temp.date, '/20', '/')
    return(as.Date(temp.date, "%d/%m/%y"))   
  }

 ## Convert event date

  # Date
  ltr.data$event.date <- fixLTDates(ltr.data$event.date)

  # Year
  ltr.data$event.year <- as.numeric(substr(ltr.data$event.date, 1, 4))

 ## Converted list data date fields

  listing.data$first.adv.date <- fixLTDates(listing.data$first.adv.date)
  listing.data$last.adv.date <- fixLTDates(listing.data$last.adv.date)

```

&nbsp;

#### Fix Boolean Variables

Next we fix a few of the boolean variables in the long term data that are stored in a non-compatible format.

```{r fix_lt_bools, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr.data$has.study <- ifelse(ltr.data$has.study == 'True', 1, 0)
  ltr.data$has.courtyard <- ifelse(ltr.data$has.courtyard == 'True', 1, 0)
  ltr.data$has.aircon <- ifelse(ltr.data$has.aircon == 'True', 1, 0)
  ltr.data$has.balcony <- ifelse(ltr.data$has.balcony == 'True', 1, 0)
  ltr.data$has.garage <- ifelse(ltr.data$has.garage == 'True', 1, 0)

```

&nbsp;

### Standardize Property Types

&nbsp;

AirBNB units are classified into 19 diffeltr property type.  The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that the Igloo, at the very least, is not a completely true representation of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

1. **House**:  Includes properties labeled 'House' or 'Townhouse'
2. **Apartment**: Includes properties labeled as 'Apartment' or 'Condominium'
3. **Other**: Includes properties labeled as 'Bed & Breakfast', 'Boat', 'Bungalow', 'Cabin', 'Camper/RV', 'Chalet', 'Dorm', 'Earth House', 'Hut', 'Igloo', 'Loft', 'Other', 'Tent', 'Treehouse', Villa' and 'Yurt.

Within the long-term ltral data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories:

1. **House**:  Includes properties labeled 'Duplex', 'House', 'Terrace' or 'Townhouse'
2. **Apartment**: Includes properties labeled as 'Unit' or 'Studio'
3. **Other**: Includes properties labeled as 'Semi' and 'Villa'

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the AirBNB data conforms to North American lexicon.  This is most appaltr when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).  Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

We then create a new variable in each dataset and give each one of the three categories discussed above.

```{r fix_type, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Short Term data

  abb.data$type <- 'Other'
  abb.data$type[abb.data$property.type %in% c('Apartment', 'Condominium')] <- 'Apartment'
  abb.data$type[abb.data$property.type %in% c('House', 'Townhouse')] <- 'House'

 ## Long Term data

  ltr.data$type <- 'Other'
  ltr.data$type[ltr.data$property.type %in% c('Duplex', 'House',
                                           'Terrace', 'Townhouse')] <- 'House'
  ltr.data$type[ltr.data$property.type %in% c('Studio', 'Unit')] <- 'Apartment'

```

The **Other** categories contain very small percentage of properties in both the AirBNB and the ltral data.  As a result, we remove the 'Others' from the data.

```{r remove_others, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE, fig.height=4}

  abb.data <- subset(abb.data, type != 'Other')
  ltr.data <- subset(ltr.data, type != 'Other')

```

Next, for the short term ltrals units can be listed as one of three types relating to the extent of the property which is able to be booked:

1. **Entire Home/Apt**: The entire home or apartment is available
2. **Private Room**: One room within a house or apartment is available
3. **Shared Room**: A bed within a room shared by other occupants is available

As long-term ltrals do not offer **Private Room** or **Shared Room** options and our purpose here is a comparison of the two we remove all short term properties that do not lease the entire home/apt. Unfortunately, this filter does remove about 40% of the short term data, however, given the research question it is unavoidable.

```{r ltype_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE, fig.height=4}

  abb.data$listing.type <- as.character(abb.data$listing.type)
  abb.data$listing.type[abb.data$listing.type == 'Entire home/apt'] <- 'Entire home'

  ggplot(abb.data, aes(x=listing.type)) + 
    geom_bar(color=abb.col[1], fill=abb.col[1]) +
    xlab('\nListing Type\n') + 
    ggtitle('\nShort Term ltrals\n')  +
    ylab('# of Properties\n')

```

```{r keep_entire_home, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  abb.room <- abb.data[abb.data$listing.type != 'Entire home', ]
  abb.data <- abb.data[abb.data$listing.type == 'Entire home', ]

```

### Trim by Time

Next we trim the observations by time.  Our hypothetical simulated decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long term ltral versus an AirBNB, short term ltral approach to generating income.  To determine the likely long term ltral rate for any given property, we will use the observed ltral transactions from the market from September 1, 2014 to August 31, 2015.  Taking the AirBNB approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factor).  In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.  Another reason for using these two period of data is that the AirBNB data has missing daily values prior to September 1, 2016.  While we have imputed these missing daily observations, we hold the actual observed values to be of greater accuracy and have designed the initial simulated example to take advantage of the full observed set.

```{r clip_time, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

 ## Limit short term data

  abb.dataf <- abb.data[abb.data$min.date >= '2015-09-01', ]
  daily.dataf <- daily.data[daily.data$date >= '2015-09-01', ]

 ## Limit long term data

  ltr.dataf <- ltr.data[ltr.data$event.date >= '2014-09-01' & 
                            ltr.data$event.date <= '2015-08-31', ]
  listing.dataf <- listing.data[listing.data$id.key %in% ltr.dataf$id.key, ]

```

### Locations and Submarkets

Location is clearly an important determinant in both the short and long term ltral markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs.  In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts.  The specific suburb of a property is the second, or finer scale, at which the market operates. The plot below shows the three large submarkets.

```{r mappoints, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Convert shapefile to ggplot-able object

  # Fix lat/long

  # Limit to extent of data
  sub.table <- table(ltr.dataf$suburb)
  sub.table <- sub.table[sub.table != 0]

  # Clip to only those with a ltral observation
  study.suburbs <- suburbs.shp[suburbs.shp@data$name_2006 %in% names(sub.table),]

  # Convert to an outer bound with holes
  ss.bound <- gUnaryUnion(study.suburbs)

  # Remove holes from the shapefile
  ssb <- slot(ss.bound, "polygons") 
  holes <- lapply(ssb, function(x) sapply(slot(x, "Polygons"), 
                                          slot, "hole"))
  res <- lapply(1:length(ssb), function(i) slot(ssb[[i]], "Polygons")[!holes[[i]]]) 
  IDs <- row.names(ss.bound) 
  ssb.fill <- SpatialPolygons(lapply(1:length(res), function(i) 
   Polygons(res[[i]], ID=IDs[i])), proj4string=CRS(proj4string(ss.bound)))

  # Identify which suburbs all in area covered by outer bound
  ss.in <- gIntersects(suburbs.shp, ssb.fill, byid=T)
  study.suburbs <- suburbs.shp[which(ss.in), ]

  # Convert into a ggplot-able format
  subs <- fortify(study.suburbs, region='name_2006')

  # Add suburb category
  subs$des <- suburbs.shp@data$suburbdesi[
    match(subs$id, suburbs.shp@data$name_2006)]

```

```{r mappoints2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.height=5}

 ## Build the map object

  # Set colors
  g.cols <- abb.col[c(2,10,9)]

  # Make map
  loc.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group, fill=des), alpha=.8) + 
    scale_fill_manual(values=g.cols, labels=c('Inner      ',
                                              'Middle      ',
                                              'Outer')) +
    geom_path(data=subs, aes(x=long, y=lat, group=group), col='gray40') +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    ggtitle("Melbourne Suburb Classifications")

## Plot map

  loc.map

```

It is likely that the short-term ltral market follows a similar market hierarchy, however, we can imagine that a few diffeltr spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beach; and 2) Proximity to key tourist activities and events.  Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analysing short and long term ltrals in Melboure:

1. Rural (Outer suburbs, not Beach)
2. Suburban (Middle suburbs, not Beach)
3. City (Inner Suburbs, not Beach, not Core)
4. City-Core (Select inner suburbs with tourist activities and near CBD)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short and long term data.  We also add the Statistical Area 1 code for each property for use in additional spatial modeling later on. Fainlly, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs).

```{r core_suburbs, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 core.suburbs <- c('Albert Park', 'Carlton', 'Collingwood', 'Cremorne', 'Docklands',
                   'East Melbourne', 'Fitzroy', 'Melbourne', 'Port Melbourne', 'Prahran',
                   'Richmond', 'South Melbourne', 'South Yarra', 'Southbank', 'St Kilda',
                   'St Kilda West', 'West Melbourne', 'Windsor')

```

```{r beachbuffer, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Determine centroid of study area

  latMid <- median(abb.dataf$latitude)
  longMid <- median(abb.dataf$longitude)

 ## Calculate Degress to Meters conversion

  m.lat = 111132.954 - 559.822 * cos(2 * latMid) + 1.175 * cos(4 * latMid)
  m.long = 111132.954 * cos (latMid)
  m.conv <- mean(m.lat, m.long)

 ## Create the beach buffer

  beach.buffer <- gBuffer(beach.shp, byid=FALSE, width=500/m.conv)

```

```{r add_st_space, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Short term data

  # Convert to point shapefile
  abb.prop.shp <- SpatialPointsDataFrame(cbind(abb.dataf$longitude,
                                               abb.dataf$latitude),
                                         abb.dataf,
                                         proj4string=CRS("+init=epsg:4283"))

  # Add suburb name 
  spJoin <- over(abb.prop.shp, suburbs.shp)
  abb.dataf$suburb <- as.character(spJoin$name_2006)
  abb.dataf$suburb[is.na(abb.dataf$suburb)] <- 'missing'

  # Add suburb designation
  abb.dataf$sub.class <- as.character(spJoin$suburbdesi)
  abb.dataf$sub.class[is.na(abb.dataf$sub.class)] <- 'missing'

  # Add SA1 designation
  spJoin <- over(abb.prop.shp, sa1s.shp)
  abb.dataf$sa1 <- spJoin$sa1_main11

  ## Add Beach Designation
  spJoin <- over(abb.prop.shp, beach.buffer)
  abb.dataf$beach <- ifelse(is.na(spJoin), 0 ,1)

  # Remove those missing a suburb classification (out of area)
  abb.dataf <- abb.dataf[abb.dataf$sub.class != 'missing', ]
  abb.dataf <- abb.dataf[abb.dataf$sa1 != 'missing', ]

```

```{r convert_st_subm, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Add the submarket designations

  # Add individual classes
  abb.dataf$sub.mrkt <- 'rural'
  abb.dataf$sub.mrkt[abb.dataf$sub.class == 'Middle'] <- 'suburban'
  abb.dataf$sub.mrkt[abb.dataf$sub.class == 'Inner'] <- 'city'
  abb.dataf$sub.mrkt[abb.dataf$suburb %in% core.suburbs] <- 'city-core'
  abb.dataf$sub.mrkt[abb.dataf$beach == 1] <- 'beach'

  # Order the factors properly
  abb.dataf$sub.mrkt <- factor(abb.dataf$sub.mrkt, 
                               levels=c('city-core', 'city', 'suburban', 
                                        'rural', 'beach'))

```

```{r abb.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Make map
  abb.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group), color='gray50') +
    geom_point(data=abb.dataf, aes(x=longitude, y=latitude, color=sub.mrkt), size=.2) +
    scale_color_manual(values=sm.col) +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3))) + 
    ggtitle('AirBNB Location\nBy Submarket')

```

About 5% of the long term ltral observations have missing latitude and longitude values.  Half of these do, however, have a lat/long value for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps for those properties that lack a specific lat/long value we apply the street centroid value.  For those without either (3% of the total) we remove them from the dataset at this point.  After removing these, we apply the submarket designations.

```{r fix_ltt_geo, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Label those missing lat/long

  miss.LL <- which(is.na(ltr.dataf$latitude))

 ## Create a field to indicate the approximate matches

  ltr.dataf$miss.LL <- 0
  ltr.dataf$miss.LL[miss.LL] <- 1

 ## Apply approximate match

  ltr.dataf$latitude[miss.LL] <- ltr.dataf$street.latitude[miss.LL]
  ltr.dataf$longitude[miss.LL] <- ltr.dataf$street.longitude[miss.LL]

 ## Remove those still missing latitude data

  ltr.dataf <- ltr.dataf[!is.na(ltr.dataf$latitude), ]

```

```{r add_lt_space, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Long term data

  # Convert to point shapefile
  ltr.prop.shp <- SpatialPointsDataFrame(cbind(ltr.dataf$longitude,
                                                ltr.dataf$latitude),
                                         ltr.dataf,
                                         proj4string=CRS("+init=epsg:4283"))

  # Add suburb name 
  spJoin <- over(ltr.prop.shp, suburbs.shp)
  ltr.dataf$suburb <- as.character(spJoin$name_2006)
  ltr.dataf$suburb[is.na(ltr.dataf$suburb)] <- 'missing'

  # Add suburb designation
  ltr.dataf$sub.class <- as.character(spJoin$suburbdesi)
  ltr.dataf$sub.class[is.na(ltr.dataf$sub.class)] <- 'missing'

  # Add SA1 designation
  spJoin <- over(ltr.prop.shp, sa1s.shp)
  ltr.dataf$sa1 <- spJoin$sa1_main11

  ## Add Beach Designation
  spJoin <- over(ltr.prop.shp, beach.buffer)
  ltr.dataf$beach <- ifelse(is.na(spJoin), 0 ,1)

  # Remove those missing a suburb classification (out of area)
  ltr.dataf <- ltr.dataf[ltr.dataf$sub.class != 'missing', ]
  ltr.dataf <- ltr.dataf[ltr.dataf$sa1 != 'missing', ]

```

```{r convert_lt_subm, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Add the submarket designations

  # Add individual classes
  ltr.dataf$sub.mrkt <- 'rural'
  ltr.dataf$sub.mrkt[ltr.dataf$sub.class == 'Middle'] <- 'suburban'
  ltr.dataf$sub.mrkt[ltr.dataf$sub.class == 'Inner'] <- 'city'
  ltr.dataf$sub.mrkt[ltr.dataf$suburb %in% core.suburbs] <- 'city-core'
  ltr.dataf$sub.mrkt[ltr.dataf$beach == 1] <- 'beach'

  # Order the factors properly
  ltr.dataf$sub.mrkt <- factor(ltr.dataf$sub.mrkt, 
                               levels=c('city-core', 'city', 'suburban', 
                                        'rural', 'beach'))

```

```{r ltr.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Make map
  ltr.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group), color='gray50') +
    geom_point(data=ltr.dataf, aes(x=longitude, y=latitude, color=sub.mrkt), size=.1) +
    scale_color_manual(values=sm.col) +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3))) + 
    ggtitle('Long Term ltral Location\nBy Submarket')

```

```{r loc.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE, fig.width=9}

  ggMultiPlots(abb.map, ltr.map, cols=2)

```

```{r smtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE, fig.width=9}

  # Make map
  abbtype.map <- ggplot() +
    geom_bar(data=abb.dataf, aes(x=sub.mrkt, fill=sub.mrkt)) +
    scale_fill_manual(values=sm.col) +
    facet_wrap(~type) +
    ylab('# of Properties\n') + xlab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    ggtitle('AirBNB Property Type\nBy Submarket')

  # Make map
  ltrtype.map <- ggplot() +
    geom_bar(data=ltr.dataf, aes(x=sub.mrkt, fill=sub.mrkt)) +
    scale_fill_manual(values=sm.col) +
    facet_wrap(~type) +
    ylab('# of Properties\n') + xlab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    ggtitle('Long Term Property Type\nBy Submarket')

```

```{r loctype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE, fig.width=9}

  ggMultiPlots(abbtype.map, ltrtype.map, cols=2)

```

## By Values (outliers, etc.)

We now filter both datasets by removing observed values unlikely to represent 'typical' short or long term ltral properties.

### Bedrooms

We start by looking at bedrooms. We begin by removing all properties that are missing bedrooms information or have a likely data error in this fields (more than 14 bedrooms). We then plot the distribution of diffeltr bedrooms counts.  From this point onwards we bifurcate our analysis between houses and apartments as the markets (prices, rates, and properties) differ between the two property types.

```{r bed_fltr, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Remove missing and data errors

  # Long term data
  ltr.dataf <- ltr.dataf[!is.na(ltr.dataf$bedrooms), ]
  ltr.dataf <- ltr.dataf[ltr.dataf$bedrooms <= 10, ]

  # Short term data
  abb.dataf <- abb.dataf[!is.na(abb.dataf$bedrooms), ]
  abb.dataf <- abb.dataf[abb.dataf$bedrooms <= 10, ]

```

```{r bed_fltr2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.height=5}

 ## Build long term bar plot

  ltrbed <- ggplot(ltr.dataf, aes(x=bedrooms)) + 
    geom_bar(fill=abb.col[3], color=abb.col[3]) +
    facet_wrap(~type) +
    scale_x_continuous(labels=1:10, breaks=1:10)+
    xlab('\nBedrooms\n') +
    ylab('') +
    ggtitle('Bedrooms in Long Term ltrals')

 # Build short term bar plot

  abbbed <- ggplot(abb.dataf, aes(x=bedrooms)) + 
    geom_bar(fill=abb.col[1], color=abb.col[1]) +
    facet_wrap(~type) +
    scale_x_continuous(labels=0:10, breaks=0:10)+
    xlab('\nBedrooms\n') +
    ylab('') +
    ggtitle('Bedrooms in Short Term ltrals')

 ## Make the comparison plot

  ggMultiPlots(abbbed, ltrbed, cols=1)

```

For both short and long term ltrals we see that most apartments have less than 4 bedrooms and houses less than 5.  We filter the data accordingly. Additionally, note that short term ltrals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units.

```{r bed_fltr3, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Remove all with more than 4 beds

  abb.dataf <- abb.dataf[abb.dataf$bedrooms < 5, ]
  ltr.dataf <- ltr.dataf[ltr.dataf$bedrooms < 5, ]

 ## Remove apartments with more than 3

  abba <- which(abb.dataf$bedrooms > 3 & abb.dataf$type =='Apartment')
  ltra <- which(ltr.dataf$bedrooms > 3 & ltr.dataf$type =='Apartment')
  if(length(abba)>0) abb.dataf <- abb.dataf[-abba, ]
  if(length(ltra)>0) ltr.dataf <- ltr.dataf[-ltra, ]

  # Convert 0 Bed to 1 Bed  
  abb.dataf$bedrooms[abb.dataf$bedrooms == 0] <- 1

```

### Bathrooms

We then turn to bathrooms, starting by looking at the distribution of baths between short/long and apt/house.

```{r bath_fltr, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Remove missing data
ltr.dataf <- ltr.dataf[!is.na(ltr.dataf$baths), ]
abb.dataf <- abb.dataf[!is.na(abb.dataf$baths), ]

```

```{r bath_fltr2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.height=5}

 ## Build long term plot

  ltbath <- ggplot(ltr.dataf, aes(x=baths)) + 
    geom_bar(fill=abb.col[3], color=abb.col[3]) +
    facet_wrap(~type) +
    scale_x_continuous(labels=1:8, breaks=1:8)+
    xlab('\nBathrooms\n') +
    ylab('') + 
    ggtitle('Bathrooms in Long Term ltrals')

 ## Build short term plot

  stbath <- ggplot(abb.dataf, aes(x=baths)) + 
    geom_bar(fill=abb.col[1], color=abb.col[1]) +
    facet_wrap(~type) +
    scale_x_continuous(labels=0:8, breaks=0:8)+
    xlab('\nBathrooms\n') +
    ylab('') + 
    ggtitle('Bathrooms in Short Term ltrals')

 ## Make Plots

  ggMultiPlots(stbath, ltbath, cols=1)

```

Here we see that nearly all properties have 3 or fewer bathrooms and we filter the data accordingly. A few short term properties state no baths, which is likely an error so we remove these as well. Also, here we notice that the short term data gives bathrooms in halves, while the long term data presents baths in whole numbers.  We round the baths up to whole integers.

```{r bath_fltr3, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Filter by bath count

  # Short term data
  abb.dataf <- abb.dataf[abb.dataf$baths < 4, ]
  abb.dataf <- abb.dataf[abb.dataf$baths > 0, ]

  # Long term data
  ltr.dataf <- ltr.dataf[ltr.dataf$baths < 4, ]

 ## Round baths up

  abb.dataf$baths <- round(abb.dataf$baths + .01, 0)

```

### Bed/Bath Combination

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short term ltral) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two.

```{r filtBB, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Create combination variable

  ltr.dataf$bedbath <- paste0(ltr.dataf$bedrooms, '..', ltr.dataf$baths)
  abb.dataf$bedbath <- paste0(abb.dataf$bedrooms, '..', abb.dataf$baths)

```

```{r filtBB2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.height=5}

 ## Make long term plot

  ltb <- ggplot(ltr.dataf, aes(x=bedbath)) + 
    geom_bar(fill=abb.col[3], color=abb.col[3]) +
    xlab('\nBed/Bath Combo\n') +
    ylab('Number of Properties\n') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle('Long Term ltrals') + 
    facet_wrap(~type)

 ## Make short term plots

  stb <- ggplot(abb.dataf, aes(x=bedbath)) +
    geom_bar(fill=abb.col[1], color=abb.col[1]) + 
    xlab('\nBed/Bath Combo\n') +
    ylab('Number of Properties\n') +
    ggtitle('Short Term ltrals') + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap(~type)

 ## Plot

  ggMultiPlots(ltb, stb, cols=1)

```

From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2 and 4/2. We filter the data to these combinations.

```{r filtBB3, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Set level of acceptable combinations

  acc.bb <- c('1..1', '2..1', '2..2', '3..1', '3..2', '4..2')

 ## Filter data

  ltr.dataf <- ltr.dataf[ltr.dataf$bedbath %in% acc.bb, ]
  abb.dataf <- abb.dataf[abb.dataf$bedbath %in% acc.bb, ]

```

### Rates and ltrs

Finally, we turn to filtering by the prices -- the nightly rates and weekly ltral values.  We start by removing observations missing price data.

```{r ltprice_fltr, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr.dataf <- ltr.dataf[!is.na(ltr.dataf$event.price), ]
  abb.dataf <- abb.dataf[!is.na(abb.dataf$nightly.rate), ]

```

```{r ltprice_fltr2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  # ltr Plot
  ltr.plot <- ggplot(ltr.dataf, aes(x=event.price)) + 
    geom_density(fill=abb.col[3],alpha=.8) +
    ylab('') +
    xlab('Weekly ltr') + 
    ggtitle('Long Term ltrals') + 
    scale_x_continuous(limits=c(0, 1500)) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

  # Nightly Rate plot
  rate.plot <- ggplot(abb.dataf, aes(x=nightly.rate)) + 
    geom_density(fill=abb.col[1], alpha=.8) + 
    ylab('') +
    xlab('Nightly Rate') +
    ggtitle('Short Term ltrals') + 
    scale_x_continuous(limits=c(0, 600)) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

```

```{r ltprice_fltr2s, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

ggMultiPlots(ltr.plot, rate.plot, cols=2)

```

From these density plots we see that most weekly ltrals are greater than $200 and less than $1000 per week while most nightly rates are greater than $50 and less than $400 per night. We filter the data accordingly.

```{r ltprice_fltr3, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Filter the long term data

  ltr.dataf <- ltr.dataf[ltr.dataf$event.price >= 200 & 
                         ltr.dataf$event.price <= 1000, ]

 ## Filter the short term data

  abb.dataf <- abb.dataf[abb.dataf$nightly.rate >= 50 &
                         abb.dataf$nightly.rate <= 400, ]

```

## Filter Transaction Data

The previous exercises filtered the property-level data.  We now filter the transaction level datasets so that they only include those property that remain in the filtered property-level datasets

```{r filttransdata, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Short term data

  daily.dataf <- daily.dataf[daily.dataf$property.id %in% abb.dataf$property.id,]

 ## Long term list data

  listing.dataf <- listing.dataf[listing.dataf$activity.id %in% ltr.dataf$activity.id,]

```

Finally, we convert the prepared data back to the original object names.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  abb.data <- abb.dataf
  daily.data <- daily.dataf
  ltr.data <- ltr.dataf
  listing.data <- listing.dataf

```

# Prepare for Revenue Estimation

We now turn to calculating the revenues for each property under an Airbnb and a long term rental scenario.  Before we start, there are few additional data prepartion tasks to undertake.

First we calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed for the Airbnb properties only.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Determine # of booking for each airbnb
  abb.data$bookings <- (abb.data$total.days - 
                          abb.data$blocked.days - 
                           abb.data$available.days)

  # Determine property specific occupancy rate
  abb.data$occ.rate <- (abb.data$bookings / 
                           (abb.data$total.days - abb.data$blocked.days))

  # Set the blocked rate
  abb.data$blocked.rate <- abb.data$blocked.days / abb.data$total.days

  # Set the extrapolation parameter for those on the site for less than whole period
  abb.data$extr.par <- 366 / abb.data$total.days

```

Next we extract all of the daily reservation data that fits our criteria -- reserved, not imputed and not over $600 per night. From this information we calculate the median nightly reservation price for each property and append it to the property level data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Get all of the reservations
  resv.daily <- daily.data[daily.data$status == 'R' &
                             daily.data$booked.date != 'imputed' &
                              daily.data$price < 600, ]

  # Calculate the property specific median nightly rate
  med.daily <- tapply2DF(resv.daily$price, resv.daily$property.id, median)

  # Add median rate to the property  
  abb.data$med.rate <- med.daily$Var[match(abb.data$property.id, med.daily$ID)]

```

For the airbnb properties we then calculate the actual revenue over the Sep 1 2015 to Aug 31 2016 period. Those with no revenue are removed.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Calculate the total revenue  
  abb.data$revenue <- (abb.data$med.rate * exch.rate * 
                          abb.data$extr.par * abb.data$bookings)

  # Remove those with no revenue
  abb.data <- abb.data[!is.na(abb.data$revenue), ]

```

Next, we calculate the revenue for the long term rental data, saving only those with revenue of at least $5,000 (less than this represented properties with very long days on market.)

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 ## Calculate ltral revenues

  # Calculate the actual revenue
  ltr.data$revenue <- ltr.data$event.price * (52 - ltr.data$dom / 7)

  # Remove those with very low revenues (due to very long DOMs)
  ltr.data <- ltr.data[ltr.data$revenue > 5000, ]

 ## Save for future costs

```

Finally, we create and fix three variables. One factor variable for bed/bath combination, a factor variable that splits bed/bath by house and apartment and then we convert suburb from a factor to a character.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Re-order factors
  abb.data$bedbath <- factor(abb.data$bedbath, 
                              levels=c('2..2', '1..1', '2..1', '3..1', '3..2', '4..2'))
  ltr.data$bedbath <- factor(ltr.data$bedbath, 
                               levels=c('2..2', '1..1', '2..1', '3..1', '3..2', '4..2'))

  # Add product as a variable
  abb.data$product <- paste0(substr(abb.data$type, 1, 1), abb.data$bedbath)
  ltr.data$product <- paste0(substr(ltr.data$type, 1, 1), ltr.data$bedbath)

  # Convert suburbs from factor to character
  abb.data$suburb <- as.character(abb.data$suburb)
  ltr.data$suburb <- as.character(ltr.data$suburb)

```

### Save Workspace for using in 'Working' script

```{r savewrkspc, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  save.image("C:/Dropbox/Research/airBNB/data/analyzed/abb_working.RData")

```

## Cross impute rates and ltrs

### Global

### By submarkets

### Matching for validation

### - Short to Long

### - Long to Short

## Vacancy Models

# Estimate the ltral income for all properties

## Other costs/fees/income considerations

## Overall rate of return calculations and comparisons

### - Rates by Product Type (Bed/Bath)

### - Rates by Location (Suburb)

### - Rates by Proximity (Near train stations / beach / event grounds, etc.)