### 2017-01-20 abb_ExecSumm.Rmd
add exec summary

---
title: "Should I Airbnb my Property?:  \n Executive Summary"
author: "Andy Krause and Gideon Aschwanden"
date: "January 16, 2017"
output: pdf_document
---

This research sets out to test under which conditions, if any, a short-term leasing strategy (Airbnb) generates more revenue than a traditional long-term approach.  Our data and analysis focus on the Melbourne metropolitan region.  To make comparisons between short and long-term markets we analyze observed and estimated revenues for a particular scenario where a property owner purchases a property on September 1 2015 and makes a decision between long or short-term rental for the next 12 months.

Our exploratory analysis shows that the univariate distribution of actual short-term revenues (from Airbnb) is much more variable than observed long term revenue.  Most Airbnb'd properties makes less but some make more, much more. To form a better comparison of revenues between like properties, we impute likely long-term revenues for the short-term properties using a model based on the long-term rental market observations. Over the entire market, we find that 31\% of the Airbnb properties did collect more revenue than they would likely do so with a long-term leasing strategy.  The probability of preference for a short-term strategy, however, is not equal for all types of properties.  More specifically, we find that short-term preference is higher in:

1. Apartments versus homes
2. Larger properties versus smaller properties
3. Properties in the city core (central 18 suburbs)
4. Rural areas as opposed to suburban (but not versus city properties)

We also find the following host behaviors or policies to result in higher probability of short-term preference:

1. More allowed guests per bedroom
2. Shorter minimum stays
3. More strict cancellation policies

Overall, we find that, indeed, some properties may generate more revenue under a short-term approach than in a long-term strategy. The probability of acheiving short-term preference is highest for large apartments in the city core with low minimum stays, strict cancellation policies and a high number of guests per bedroom. However, even within the core city apartment market an owner must achieve a nightly rate and an overall occupancy rate both above the median in order to reach an equal probability of short-term preference.  In other words, there is solid competition in this market and short-term preference depends on proper pricing and high occupancy.

It is no surprise, given these results, that inner city apartment make up a large proportion of the Airbnb market in Melbourne.  The market appears aware, to some extent, of the profitability of this property type under a short-term approach.