### 2017-10-25 abbDataAnalysis_Script.Rmd

---
title: "Should I Airbnb My Property: Analytic Code Explanation"
author: "Andy Krause and Gideon Aschwanden"
date: "March 10, 2017"
output: html_document
---

### Introduction

This document describes all of the code used to the generate the output included in the 'Should I Airbnb my Property' study.  Additional analysis that were not included in the final report are also found within this document.  This analysis uses the output from the **abbDataClean_Script.rmd** file found in the same online repository. 
### Preliminary Commands

We begin by loading the libraries necessary to do this analysis.

&nbsp;

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)
  library(kernlab)
  library(lmtest)
  library(ROCR)

```

&nbsp;

```{r set_dates}

  data_name <- 'jun'
  strike_date <- lubridate::as_date('2015-06-01')

```

We then set the paths to the data, code and figures, based on the user of the document.  We also set a variable indicating whether or not we are working offline.

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

    data.path <- 'c:/dropbox/research/airBNB/data/'
    code.path <- 'c:/code/'
    fig.path <- 'c:/dropbox/research/airBNB/figures/'

```

Next, we source the necessary custom code files for the analysis.  We also set a few options here.

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

 # Current Project files

  source(paste0(code.path, "research/AirBNBMelbourne/abb_Functions.R"))

 # Other custom files

  if(offline){

  source(paste0(code.path, 'tools/dataVizTools/ggPlotTools.R'))
  source(paste0(code.path, 'tools/datamgmttools/dataMungeTools.R'))
  source(paste0(code.path, 'tools/spatialTools/spatialLogitModels.R'))
  source(paste0(code.path, 'tools/spatialTools/point2surface_Tools.R'))

  } else {

  source(paste0('https://raw.githubusercontent.com/andykrause/dataVizTools',
                '/master/ggPlotTools.R'))
  source(paste0('https://raw.githubusercontent.com/andykrause/datamgmttools',
                '/master/dataMungeTools.R'))
  source(paste0('https://raw.githubusercontent.com/andykrause/spatialTools',
                '/master/spatialLogitModels.R'))
  source(paste0('https://raw.githubusercontent.com/andykrause/spatialTools',
                '/master/point2surface_Tools.R'))

  }

 ## Set options

  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

```

We load the workspace from the data clean script that includes the prepared data objects.

```{r load_str_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Property Information

  load(file.path(data.path, 'prepared', paste0('prep_workspace_', data_name, '.RData')))

```

### Calculating Net Revenues

We begin the analysis by calculating the net revenues for the short term data.  We calculate three types of net revenues, Observed, Extrapolated (Actual) and Potential. The difference are described more fully in the writeup.

Before we start, we estimate two figures necessary to determine utility costs: 1) Costs per person per day (cpppd) and the average number of guests per bedroom.The State of Victoria gives energy costs at \$330 (apartment) to \$520 (house).  For 2 persons in an apartment that works out to \$5 per person per day.  For a house of 4 persons that works out to about \$4 per person per day. We use a mid point value of \$4.50 per person per day.  We use an estimate of 1.5 persons per bedroom in each unit.

```{r utility_estimates, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_df$cpppd <- 4.5
  str_df$avg.guests <- str_df$bedrooms * 1.5

```

```{r obs_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_df$str.obs.revenue <- ((str_df$bookings * str_df$med.rate) - 
                                 (str_df$bookings * str_df$avg.guests *
                                    str_df$cpppd))

```

```{r act_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_df$extr.par <- 366 / str_df$total.days

  str_df$str.act.revenue <- ((str_df$bookings * str_df$extr.par *
                                  str_df$med.rate) - 
                                 (str_df$bookings * str_df$avg.guests * 
                                    str_df$cpppd * str_df$extr.par))
```

```{r pot_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_df$occ.rate <- 1 - (str_df$block.rate + str_df$avail.rate)

  str_df$pot.occ.rate <- ((str_df$bookings * str_df$extr.par) / 
    (366 - (366 * str_df$block.rate)))

  str_df$str.pot.revenue <- ((str_df$pot.occ.rate * 366 * str_df$med.rate) - 
                                 (str_df$pot.occ.rate * 366 * 
                                    str_df$avg.guests * str_df$cpppd))

```

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr.data$ltr.revenue <- ltr.data$gross.revenue

```

#### Compare Revenue Distributions

Before we examine the distributions of the observed short-term revenues and the estimated long-term revenues, we must produce the estimates of long-term revenue for the short-term properties.

We start by specifying the imputation model specification we will use to estimated likely long-term (weekly) rental prices for each short-term property.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr.mod.spec <- formula(log(event.price) ~ as.factor(type) + 
                            as.factor(bedbath) + as.factor(suburb) + 
                            as.factor(ltr.month))

```

We then employ a custom function to impute the long term rents for the short-term properties.  The 'clip.field' is a spatial fixed effects that add spatial recognition to the models.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  imp.data <- imputeLtrRents(ltr.df=ltr.data,
                               str.df=str_df,
                               mod.spec=ltr.mod.spec,
                               match.factor='suburb')

```

From the imputed results, we extract the imputed rental values and merge to the original short-term data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Extract out DFs (only those that imputed)
  str.metro <- merge(str_df, imp.data$imp.rent, by='property.id')

```

Next we impute the likely long-term search cost (days on market) for the short-term properties.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Apply dom to abb
  str.metro$imp.dom <- imputeDOM(str.metro, ltr.data, calc.type='median')

```

With the imputed DOMs and weekly rental rates we now make the an estimate of likely annual revenue if leased on the long term market. We append this to the original dataset.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str.metro$ltr.imp.revenue <- str.metro$imp.rent * (52 - str.metro$imp.dom / 7)

```

Finally we make a comparison of the observed short term revenues and the imputed long term revenue and add these field back onto the original data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  comp.revs <- compareRevenues(str.metro)

  str.metro <- merge(str.metro, comp.revs, by='property.id')

```

The above process of long-term revenue estimation treated the entire market as operating similarly. Realizing that the imputed long term rental values and search times may be more accurate when estimated at smaller scales, we functionalize the above processes and then apply it to short-term rentals, broken into submarkets by type, bed/bath and geographic area.  In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all compbinations of type, bed/bath and geographic area. We then combine the results form all estimates back into a single data.frame that contains all of the short-term rentals. We use these finer scale estimates going forward (the data in the **str_df.ic** object).

```{r build.tbg.data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Create the type, bed/bath and geo sub market variable

 str_df$tbg <- paste0(str_df$product,'.', str_df$geo.mrkt)
 ltr.data$tbg <- paste0(ltr.data$product,'.', ltr.data$geo.mrkt)

 # Make IC estimates at TBG level

 str_df.ic <- abbImputeCompare(str.df=str_df,
                                 ltr.df=ltr.data,
                                 mod.spec=ltr.mod.spec,
                                 match.factor='suburb',
                                 split.field='tbg',
                                 verbose=FALSE)

```

Having calculated the revenues, both short-term and long-term imputed, we now compare the actual (extr), potential and imputed long-term revenue distributions for the short term properties

```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5}

  # Actual
   stra.rev <- str_df.ic[,c('property.id', 'str.act.revenue')]
   stra.rev$tenure <- 'Short-Term (Actual)   '
   names(stra.rev)[2] <- 'revenue'

  # Potential
   strp.rev <- str_df.ic[,c('property.id', 'str.pot.revenue')]
   strp.rev$tenure <- 'Short-Term (Potential)   '
   names(strp.rev)[2] <- 'revenue'

  # Long-Term
   ltr.rev <- str_df.ic[,c('property.id', 'ltr.imp.revenue')]
   ltr.rev$tenure <- 'Long-Term   '
   names(ltr.rev)[2] <- 'revenue'

  # Combine
   revdens.data <- rbind(stra.rev, strp.rev, ltr.rev)

  ## Build a plot

   rd.plot <- ggplot(revdens.data, 
                     aes(x=revenue, fill=tenure, color=tenure)) +
     geom_density(alpha=.5) +
     scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) +
     scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) +
     xlab('\nAnnual Revenue') +
     scale_x_continuous(breaks=c(seq(0, 75000, by=25000)),
                        labels=c('$0', '$25k', '$50k', '$75k'))+
     theme(legend.position='none',
           legend.title = element_blank(),
           plot.title = element_text(hjust = 0.5),
           axis.text.y = element_blank(),
           axis.ticks.y = element_blank(),
           axis.title.y=element_blank()) +
     coord_cartesian(xlim=c(0, 85000))
     rd.plot <- rd.plot + facet_wrap(~tenure)

  ## Plot

     rd.plot

```

## Results

We will begin by analyzing the above estimates at the entire metro scale. Analyses of various submarkets will follow. Overall, we see that 31.1\% of properties raised more revenue via short term rental than expected in the long term market.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  table(str_df.ic$str.act.pref) / nrow(str_df.ic)

```

Next, we plot the \% of properties that have short-term preference versus the occupancy rate as occupancy rate is an important determinant of short-term preference.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

 ## Set Preference data field

  str_df.ic$pref <- str_df.ic$str.act.pref

 ## Calculate the percentile fields

  str_df.ic$occ.qtl <- makeWtdQtl(str_df.ic$occ.rate, 
                                      return.type='rank') 
  str_df.ic$rate.qtl <- makeWtdQtl(str_df.ic$med.rate, 
                                  return.type='rank')

 ## Calculate raw and percentile occupancies

  rawocc.metro <- abbPrefPlot(str_df.ic,
                         x.field='occ.rate',
                         smooth=TRUE,
                         smooth.span=.5)

  qtlocc.metro <- abbPrefPlot(str_df.ic,
                        x.field='occ.qtl',
                        smooth=TRUE,
                        smooth.span=.5)

   ## Make plots

  # Raw
  raw.occ.metro <- ggplot(rawocc.metro$data,
                   aes(x=x.var, y=Var))+
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none') + 
    ggtitle('Raw Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

  # Quantile
  qtl.occ.metro <- ggplot(qtlocc.metro$data,
                   aes(x=x.var, y=Var))+
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

  # Make Plot
  ggMultiPlots(raw.occ.metro, qtl.occ.metro, cols=2)

```

Next we build a heat map that shows which combinations of occupancy rate and nightly rate are most likely to result in short-term preference.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  rate.hm <- abbHeatMap(str_df.ic,
                x.field='occ.rate',
                y.field='med.rate',
                pref.field='str.act.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  rate.hm + ylab('Nightly Rate')

```

To make the distinction more clear, we employ a Support Vector Machine to classify the areas into short and long term preference.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  rate.hm.svm <- abbHeatMap(str_df.ic,
                            x.field='occ.rate',
                            y.field='med.rate',
                            pref.field='str.act.pref',
                            svm=T, 
                            alpha.count=F,
                            add.points=T,
                            fill.colors=c(abb.col[1], abb.col[5]))
  rate.hm.svm+ ylab('Nightly Rate')

```

As the distributions of occupancy rate and nightly rate are not uniform, we standardize these values to a percentile and then replot to give a better sense of the relative level of occupancy rate and nightly rate required for short-term preference.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  qtl.hm <- abbHeatMap(str_df.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.act.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm+ ylab('Nightly  (Pcntl)')

```

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  qtl.hm.svm <- abbHeatMap(str_df.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm.svm+ ylab('Nightly  (Pcntl)')

```

### Submarkets

Now we move on to looking more deeply at various submarkets.  We start by analyzing the differences between houses and apartments.

First we make a comparison table of the results.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Make Basic comparison table
  pref.table.type <- abbCreateCompTable(ic.df=str_df.ic,
                                   split.field='type')

 # Limit to the extrapolated pref type
  pref.table.type <- pref.table.type[pref.table.type$rev.type == 'Actual', ]

```

We then turn this into a simple chart to see the differences.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ggplot(pref.table.type, aes(x=ID, weights=Var, fill=ID)) + 
    geom_bar() +
    scale_fill_manual(values=c('red', 'blue')) +
    xlab('') +
    ylab('% of properties where Airbnb is more profitable') +
    scale_y_continuous(breaks=c(0,.25,.5,.75,1),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none')

```

We then look at the relationship between occ rate and preference.  Before we start though we recalculate the percentiles within each submarket.

```{r build_occ_type1, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

 ## Create dataset

  type.data.ic <- str_df.ic

  # Id the types
  apt.id <- which(str_df.ic$type == 'Apartment')
  house.id <- which(str_df.ic$type == 'House')

  # Recalculate the percentiles
  type.data.ic$occ.qtl[apt.id] <- makeWtdQtl(type.data.ic$occ.rate[apt.id],
                                      return.type='rank')
  type.data.ic$occ.qtl[house.id] <- makeWtdQtl(
    type.data.ic$occ.rate[house.id], return.type='rank')

  type.data.ic$rate.qtl[apt.id] <- makeWtdQtl(type.data.ic$med.rate[apt.id],
                                  return.type='rank')
  type.data.ic$rate.qtl[house.id] <- makeWtdQtl(
    type.data.ic$med.rate[house.id], return.type='rank')

```

Then we calculate the preference figures

```{r build_occ_type, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

 ## Calc the occ rate numbers

  rawocc.type <- abbPrefPlot(type.data.ic,
                         x.field='occ.rate',
                         split.field='type',
                         smooth=TRUE,
                         smooth.span=.5 ,
                         spl.col=c('black', 'grey40'))

  pctocc.type <- abbPrefPlot(type.data.ic,
                         x.field='occ.qtl',
                         split.field='type',
                         smooth=TRUE,
                         smooth.span=.5 ,
                         spl.col=c('black', 'grey40'))

```

And then plot.

```{r plot_occ_type, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{smocc}Preference vs Occ. Rates: By Type'}

 ## Build raw plot

  ptr.occ <- ggplot(rawocc.type$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate') +
    scale_color_manual(values=c('black', 'grey60'),
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Raw Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

 ## Build Percentile Plots

  ptq.occ <- ggplot(pctocc.type$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_color_manual(values=c('black', 'grey60'),
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

 ## Plot

  ggMultiPlots(ptr.occ, ptq.occ, cols=2)

```

Finally, we make percentile SVM heatmaps for both apartments and houses.

```{r plot_svm_type, fig.width=8, fig.height=3.5, echo=FALSE, fig.cap='\\label{smhm}Analysis of Percentile Rates: By Type'}

 ## Make House SVM Heatmap

  house.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'House', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5])) +
  theme(legend.position='none') +
  annotate("text", x = 50, y = 50, label = "Houses", size=9) +
  ylab('Nightly Rate (Pcntl)') +
  xlab('Occupancy Rate (Pcntl)')

 ## Make Apartment SVM Heat Map

  apt.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'Apartment', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
  annotate("text", x = 50, y = 50, label = "Apartments", size=9)+
  ylab('Nightly Rate (Pcntl)') +
  xlab('Occupancy Rate (Pcntl)')

 ## Plot

  ggMultiPlots(house.hm.svm, apt.hm.svm, cols=2)

```

#### Geographic Submarkets

We now repeat the three analyses -- table, occupancy plot, heat map -- for the five geographic submarkets.

```{r table_geo, message=FALSE, warning=FALSE, results='asis', comment=FALSE, cache=TRUE, echo=FALSE, fig.cap='Geographic Submarket Preferability'}

 # Make Basic comparison table
  pref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                   split.field='geo.mrkt')

 # Convert to wide format for output
  pref.table.wide <- dcast(pref.table, ID ~ rev.type, value.var='Var')

 # Limit to the extrapolated pref type
  pref.table <- pref.table[pref.table$rev.type == 'Actual', ]
  pref.table[ ,1] <- c('City-Core', 'City', 'Suburban', 'Rural', 'Beach')
  pref.table[ ,2] <- paste0(sprintf('%.1f', pref.table[ ,2] * 100), '%')

 # Fix names  
  names(pref.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)

 ## Make printable table

  print(xtable(pref.table[,1:2], 
               caption='Geographic Submarket Preferability\n',
               digits=1,
               align='llr'),
        include.rownames=FALSE,
        caption.placement='top')

```

```{r build_occ_geo, fig.width=8, fig.height=4, echo=FALSE}

 ## Build data

  subm.data.ic <- str_df.ic

  # Get IDs
  cc.id <- which(str_df.ic$geo.mrkt == 'city-core')
  city.id <- which(str_df.ic$geo.mrkt == 'city')
  sub.id <- which(str_df.ic$geo.mrkt == 'suburban')
  rural.id <- which(str_df.ic$geo.mrkt == 'rural')
  beach.id <- which(str_df.ic$geo.mrkt == 'beach')

  # Recalc the percentiles
  subm.data.ic$occ.qtl[cc.id] <- makeWtdQtl(subm.data.ic$occ.rate[cc.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[cc.id] <- makeWtdQtl(subm.data.ic$med.rate[cc.id],
                                  return.type='rank')

  subm.data.ic$occ.qtl[city.id] <- makeWtdQtl(
      subm.data.ic$occ.rate[city.id], return.type='rank')
  subm.data.ic$rate.qtl[city.id] <- makeWtdQtl(
    subm.data.ic$med.rate[city.id], return.type='rank')

  subm.data.ic$occ.qtl[sub.id] <- makeWtdQtl(subm.data.ic$occ.rate[sub.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[sub.id] <- makeWtdQtl(subm.data.ic$med.rate[sub.id],
                                  return.type='rank')

  subm.data.ic$occ.qtl[rural.id] <- makeWtdQtl(subm.data.ic$occ.rate[rural.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[rural.id] <- makeWtdQtl(subm.data.ic$med.rate[rural.id],
                                  return.type='rank')

  subm.data.ic$occ.qtl[beach.id] <- makeWtdQtl(subm.data.ic$occ.rate[beach.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[beach.id] <- makeWtdQtl(subm.data.ic$med.rate[beach.id],
                                  return.type='rank')

```

```{r plot_occ_geo, fig.width=9, fig.height=4, message=FALSE, warning=FALSE, echo=FALSE, fig.cap='\\label{smocc}Analysis of Percentile Rates: Geographic Submarkets'}

 ## Set colors

  sm.col <- abb.col[c(1, 3, 6, 5, 2)]

 ## Create plot data

  rawocc.geo <- abbPrefPlot(subm.data.ic,
                        x.field='occ.rate',
                        split.field='geo.mrkt',
                        smooth=TRUE,
                        smooth.span=.5 ,
                        spl.col=sm.col)

  pctocc.geo <- abbPrefPlot(subm.data.ic,
                         x.field='occ.qtl',
                         split.field='geo.mrkt',
                         smooth=TRUE,
                         smooth.span=.5 ,
                         spl.col=sm.col)

 ## Make plot objects

  smr.occ <- ggplot(rawocc.geo$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate') +
    scale_color_manual(values=sm.col,
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Raw Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

  smq.occ <- ggplot(pctocc.geo$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_color_manual(values=sm.col,
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

 ## Plot

  ggMultiPlots(smr.occ, smq.occ, cols=2)

```

```{r build_svm_geo, fig.width=8, fig.height=4, echo=FALSE, fig.cap='Analysis of Percentile Rates: Metro Level'}

  # Build the SVM Heatmaps

  core.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'city-core', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5])) +
  theme(legend.position='none') +
    ylab('') +
    xlab('') +
  annotate("text", x = 50, y = 50, label = "City-Core", size=9)

  city.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'city', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "City", size=9)

  sub.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'suburban', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "Suburban", size=9)

  rural.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'rural', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "Rural", size=9)

  beach.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'beach', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "Beach", size=9)

```

```{r plot_svm_geo, message=FALSE, warning=FALSE, fig.width=8, fig.height=8, echo=FALSE, fig.cap='\\label{smhmps}Geographic Submarket Heatmaps'}

  ggMultiPlots(core.hm.svm, city.hm.svm, sub.hm.svm,
               rural.hm.svm, beach.hm.svm, cols=2)

```

## Logit models

Now we attempt to explain the drivers of preference with a set of logistic regression models.

First, a basic model with just structural characteristics.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_df.ic$bedbath <- factor(str_df.ic$bedbath,
                                levels=c('1..1', '2..1', '2..2',
                                         '3..1', '3..2', '4..2'))

  mod.act.str <- glm(str.act.pref~type+as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_df.ic)
  mod.pot.str <- glm(str.pot.pref~type+as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_df.ic)
```

Then a model that adds in the geographic submarkets

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  mod.act.loc <- glm(str.act.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)
  mod.pot.loc <- glm(str.pot.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)
```

Finally a model that adds in host policies.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  mod.act.host <- glm(str.act.pref~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      family=binomial(link='logit'),
                      data=str_df.ic)
  mod.pot.host <- glm(str.pot.pref~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      family=binomial(link='logit'),
                      data=str_df.ic)
```

We combine all model results in to a list for easy export.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  pref.models <- list(mod.act.str, mod.act.loc, mod.act.host, 
                      mod.pot.str, mod.pot.loc, mod.pot.host)
  names(pref.models) <- c('act.str', 'act.loc', 'act.host',
                          'pot.str', 'pot.loc', 'pot.host')

```

We then also run a set of regression models where the premium (or discount) for short-term leasing is the dependent variable.  We use the same three model specifications as above.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  mod.act.str.prem <- lm(str.act.prem~type+as.factor(bedbath),
                    data=str_df.ic)
  mod.pot.str.prem <- lm(str.pot.prem~type+as.factor(bedbath),
                    data=str_df.ic)

  mod.act.loc.prem <- glm(str.act.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)
  mod.pot.loc.prem <- glm(str.pot.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)

  mod.act.host.prem <- lm(str.act.prem~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      data=str_df.ic)
  mod.pot.host.prem <- lm(str.pot.prem~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      data=str_df.ic)
```

We then package these into a list for easy export

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

    prem.models <- list(mod.act.str.prem, mod.act.loc.prem, mod.act.host.prem, 
                      mod.pot.str.prem, mod.pot.loc.prem, mod.pot.host.prem)
    names(prem.models) <- c('act.str', 'act.loc', 'act.host',
                          'pot.str', 'pot.loc', 'pot.host')

```

## Sensitivity

### Potential Revenue

To test for the sensitivity of the result to the revenue type, we re-do some of the intial data vizs.  First we build a new preference table looking at the submarkets of type and geography.

```{r pot_table, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Potential Short-Term Preference'}

 # Make Basic comparison table of type
  tpref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                    split.field='type')
 # Extract Potential Results
  tpref.table <- tpref.table[tpref.table$rev.type == 'Potential', ]

 # Make baseic comparison table of submarkets  
  spref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                    split.field='geo.mrkt')

 # Extract potential results  
  spref.table <- spref.table[spref.table$rev.type == 'Potential', ]

 # Combine and round  
  st.table <- rbind(tpref.table, spref.table)
  st.table[,2] <- paste0(sprintf('%.1f', st.table[,2]*100), '%')

 ## Fix names and print

  names(st.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(st.table[,1:2], 
               caption='Potential Revenue Preferability\n',
               digits=1, align='llr'),
               include.rownames=FALSE,
               caption.placement='top',
               hline.after=c(-1, 0, 2, 7, 7))

```

Next we recrate the SVM percentile heat maps.

```{r pot_svm, results='asis', echo=FALSE, message=FALSE, comment=FALSE}

 ## create data and re-estimate the percentiles

  pot.data.ic <- str_df.ic

  pot.data.ic$occ.qtl <- makeWtdQtl(pot.data.ic$pot.occ.rate, return.type='rank')

 ## Make the pecentile heatmap

  pqtl.hm <- abbHeatMap(pot.data.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.pot.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  pqtl.hm <- pqtl.hm + ylab('Nightly Rate (Qtl)')

 ## Make percentile SVM heatmap

  pqtl.hm.svm <- abbHeatMap(pot.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.pot.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  pqtl.hm.svm <- pqtl.hm.svm + ylab('Nightly Rate (Pcntl)')+
    xlab('Occupancy Rate (Pcntl)') +
  annotate("text", x = 50, y = 50, label = "Potential", size=9)

 ## Add annotations to previous SVM

  qtl.hm.svm.p <- qtl.hm.svm +
  annotate("text", x = 50, y = 50, label = "Actual", size=9)

```

```{r plot_pot_svm, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.height=3.5, fig.cap='\\label{pothm}Heatmaps of Short-Term Preference'}

  ggMultiPlots(qtl.hm.svm.p, pqtl.hm.svm, cols=2)

```

### Host types

Finally, we test to see if grouping by hosts types has any impact.

First we must divide hosts up.  We start by looking at block rate.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  host.block.freq <- ggplot(str_df, aes(x=block.rate)) + 
  geom_density(fill=str.col, color=str.col) +
  ylab('Frequency\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0,1,.2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Host Block Rate Frequency\n') +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

  # Make Plot
  png(file=paste0(fig.path, 'dataanys_hostblock.png'), width = 480, height = 280, 
      bg='transparent')

    host.block.freq

  dev.off()

  save(host.block.freq, file=paste0(fig.path, 'hostblockfreq.RData'))

```

We then add the number of blocked periods

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  block.rate.period <- ggplot(str_df, aes(x=block.rate, y=nbr.block*extr.par)) + 
  geom_point(fill=str.col, color=str.col, size=.2) +
  ylab('Nbr of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

  # Make Plot
  png(file=paste0(fig.path, 'dataanys_blockrateperiod.png'), width = 480, height = 280, 
      bg='transparent')

    block.rate.period

  dev.off()

  block.rate.period

  save(block.rate.period, file=paste0(fig.path, 'blockrateperiod.RData'))

```

No real pattern emerges, so we divide manually.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

str_df.ic$blockpertime <- round(str_df.ic$nbr.block * str_df.ic$extr.par, 1)

str_df.ic$host.type <- 'Unknown'
str_df.ic$host.type[str_df.ic$blockpertime > 12] <- 'Multi-Platform User'
str_df.ic$host.type[str_df.ic$block.rate <= .25] <- 'Profit Seeker'
str_df.ic$host.type[str_df.ic$block.rate >= .75] <- 'Opportunistic Sharer'
str_df.ic$host.type <- factor(str_df.ic$host.type,
                             levels=c('Profit Seeker', 'Opportunistic Sharer', 
                                      'Multi-Platform User', 'Unknown'))

host.type <- ggplot(str_df.ic, 
       aes(x=block.rate, y=blockpertime, color=host.type)) + 
  geom_point(alpha=.5, size=.8, shape=16) +
  ylab('Number of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
      ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +

    scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +

  scale_color_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9])) + 
  theme(legend.position='bottom',
          legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))+
    guides(colour = guide_legend(override.aes = list(size=3,
                                                     alpha=1)))
  # Make Plot
  png(file=paste0(fig.path, 'dataanys_hosttype.png'), width = 480, height = 280, 
      bg='transparent')

    host.type

  dev.off()

  host.type

  save(host.type, file=paste0(fig.path, 'hosttype.RData'))

```

See the counts of each group

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

host.type.bar <- ggplot(str_df.ic, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  ggtitle('Short Term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))

  # Make Plot
  png(file=paste0(fig.path, 'dataanys_hosttypebar.png'), width = 480, height = 280, 
      bg='transparent')

    host.type.bar

  dev.off()

  host.type.bar

  save(host.type.bar, file=paste0(fig.path, 'hosttypebar.RData'))

```

We then re-estimate the preference table.

```{r host_table, results='asis', message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

 # Make Basic comparison table
  apref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                    split.field='host.type')

 # Limit to the extrapolated pref type
  hpref.table <- apref.table[apref.table$rev.type == 'Actual', ]
  phpref.table <- apref.table[apref.table$rev.type == 'Potential', ]
  hpref.table <- hpref.table[,1:2]
  hpref.table[,3] <- phpref.table[,2]
  hpref.table[,2] <- paste0(sprintf('%.1f', hpref.table[,2]*100), '%')
  hpref.table[,3] <- paste0(sprintf('%.1f', hpref.table[,3]*100), '%')

  names(hpref.table) <- c('Host Type', 'STP (Actual)', 'STP (Potential)')
  options(xtable.comment = FALSE)
  print(xtable(hpref.table, 
               caption='Host Type Preferability\n',
               digits=1,
               align='llrr'),
               include.rownames=FALSE,
               caption.placement='top')

```

And re-run the logistic regression models.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ps <- which(str_df.ic$host.type == 'Profit Seeker')
  os <- which(str_df.ic$host.type == 'Opportunistic Sharer')
  mu <- which(str_df.ic$host.type == 'Multi-Platform User')
  un <- which(str_df.ic$host.type == 'Unknown')

  mod.pot.ps <- glm(str.pot.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[ps,])

  mod.pot.os <- glm(str.pot.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[os,])

  mod.pot.mu <- glm(str.pot.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[mu,])

  mod.pot.un <- glm(str.pot.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[un,])

  mod.act.ps <- glm(str.act.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[ps,])

  mod.act.os <- glm(str.act.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[os,])

  mod.act.mu <- glm(str.act.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[mu,])

  mod.act.un <- glm(str.act.pref~type+as.factor(bedbath)+
                    geo.mrkt+
                    I(max.guests/bedrooms)+min.stay+
                    I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                 family=binomial(link='logit'),
                 data=str_df.ic[un,])

host.models <- list(mod.act.ps, mod.act.os, mod.act.mu, mod.act.un,
                    mod.pot.ps, mod.pot.os, mod.pot.mu, mod.pot.un)
names(host.models) <- c('act.ps', 'act.os', 'act.mu', 'act.un',
                        'pot.ps', 'pot.os', 'pot.mu', 'pot.un')

```

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  save(clean.count, daily_df, listing.data, ltr.data, pref.table,
       pref.table.wide, revdens.data, rawocc.metro, qtlocc.metro, rawocc.type,
       pctocc.type, subm.data.ic, rawocc.geo, pctocc.geo, type.data.ic, pot.data.ic,
       str_df, str_df.ic, subs,
       core.suburbs,  exch.rate, host.models, pref.models, prem.models,
       file=file.path(data.path, 'analyzed', paste0('analyzed_workspace_', data_name, 
                                                    '.RData')))

  save(str_df.ic, subs, abb.col, file=file.path(data.path,'analyzed', 
                                                paste0('shinyinput_', data_name, 
                                                       '.RData')))

 #all.mwl.coefs, all.mwl.surfs,est.points,  s.subs, sssubs,

```

<!-- ### Moving Window Logit -->

<!-- To better capture the spatial variation in the logit model results, we set up a moving window logit model that estimates a different model at each estimation point on the map.  Before beginning, we import a new suburbs shapefile and trim it to those suburbs with Airbnb properties.  -->
<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  suburbs.shp <- readShapePoly(paste0(data.path, 'geographic/melbSuburbs.shp'), -->
<!--                               proj4string=CRS("+init=epsg:4283"), -->
<!--                               delete_null_obj=TRUE) -->

<!--  ssubs <- table(str_df.ic$suburb) -->
<!--  s.subs <- subs[subs$id %in% names(ssubs),] -->
<!--  s.suburbs <- suburbs.shp[suburbs.shp@data$NAME_2006 %in% names(ssubs),] -->
<!--  sssubs <- fortify(s.suburbs) -->

<!-- ``` -->

<!-- Next we set up the prediction points.  We drap a grid over the entire area and use those points for prediction points.  -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  # Min, Max and Range -->
<!--  xmin <- min(s.subs$long) -->
<!--  xmax <- max(s.subs$long) -->
<!--  ymin <- min(s.subs$lat) -->
<!--  ymax <- max(s.subs$lat) -->
<!--  xrange <- xmax-xmin -->
<!--  yrange <- ymax-ymin -->

<!--  # Set Scale -->
<!--  scale <- mean(xrange, yrange)/100 -->

<!--  # Create estimation points -->
<!--  est.data <- str_df.ic -->
<!--  est.points <- createGridPoints(s.suburbs, scale, T) -->

<!-- ``` -->

<!-- We then specify the logistic regression model and set the moving window bandwidth.  -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  ## Set model specification   -->

<!--  pot.mod.spec <- str.pot.pref ~ type + as.factor(bedbath) +  -->
<!--    I(max.guests / bedrooms) + min.stay +  -->
<!--    I(cancel.policy == 'Flexible') +  -->
<!--    I(cancel.policy == 'Strict') -->

<!--  act.mod.spec <- str.act.pref ~ type + as.factor(bedbath) +  -->
<!--    I(max.guests / bedrooms) + min.stay +  -->
<!--    I(cancel.policy == 'Flexible') +  -->
<!--    I(cancel.policy == 'Strict') -->

<!--  ## Set bandwidth -->

<!--  k <- 400 -->

<!-- ``` -->

<!-- The results are estimated with the custom MWL function -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  pot.results.all <- mwl(est.data,  -->
<!--                         est.points,  -->
<!--                         pot.mod.spec,  -->
<!--                         k) -->

<!--  act.results.all <- mwl(est.data,  -->
<!--                         est.points,  -->
<!--                         act.mod.spec,  -->
<!--                         k) -->

<!--   pot.results.ps <- mwl(est.data[ps, ],  -->
<!--                         est.points,  -->
<!--                         pot.mod.spec,  -->
<!--                         k) -->
<!--    pot.results.os <- mwl(est.data[os, ],  -->
<!--                         est.points,  -->
<!--                         pot.mod.spec,  -->
<!--                         k) -->

<!--    pot.results.mu <- mwl(est.data[mu, ],  -->
<!--                         est.points,  -->
<!--                         pot.mod.spec,  -->
<!--                         k) -->

<!--    pot.results.un <- mwl(est.data[un, ],  -->
<!--                         est.points,  -->
<!--                         pot.mod.spec,  -->
<!--                         k) -->

<!--    act.results.ps <- mwl(est.data[ps, ],  -->
<!--                         est.points,  -->
<!--                         act.mod.spec,  -->
<!--                         k) -->
<!--    act.results.os <- mwl(est.data[os, ],  -->
<!--                         est.points,  -->
<!--                         act.mod.spec,  -->
<!--                         k) -->

<!--    act.results.mu <- mwl(est.data[mu, ],  -->
<!--                         est.points,  -->
<!--                         act.mod.spec,  -->
<!--                         k) -->

<!--    act.results.un <- mwl(est.data[un, ],  -->
<!--                         est.points,  -->
<!--                         act.mod.spec,  -->
<!--                         k) -->

<!-- ``` -->

<!-- Coefficents are extracted and put into a data.frame -->

<!-- MAKE A FUNCTION BELOW HERE -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!-- extrMWLCoef <- function(mwl.results, -->
<!--                         est.points, -->
<!--                         names.coef){ -->

<!--  coef.names <- rownames(mwl.results[[1]]) -->

<!--  # Create a list to capture -->
<!--  coef.list <- list() -->

<!--  # Create a data.frame to capture -->
<!--  coef.df <- data.frame(lat=est.points@coords[,2], -->
<!--                        long=est.points@coords[,1]) -->

<!--  # Loop through and extract coefficients -->
<!--  for(cn in 1:length(coef.names)){ -->
<!--    coef.list[[cn]] <- lapply(mwl.results, getMWLCoef, coef=coef.names[cn]) -->
<!--    coef.df[ ,ncol(coef.df) + 1] <- unlist(coef.list[[cn]]) -->
<!--  } -->

<!--  # Add plottable names -->
<!--  colnames(coef.df)[3:13] <- names.coef -->

<!--  return(coef.df) -->

<!-- } -->
<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!-- x.names <- c('intercept', 'house', 'bb11', 'bb21', 'bb31', -->
<!--                               'bb32', 'bb42', 'guest', 'minstay', 'flex', 'strict') -->

<!-- pra.coef <- extrMWLCoef(pot.results.all, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- era.coef <- extrMWLCoef(act.results.all, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- prp.coef <- extrMWLCoef(pot.results.ps, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- pro.coef <- extrMWLCoef(pot.results.os, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- prm.coef <- extrMWLCoef(pot.results.mu, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- pru.coef <- extrMWLCoef(pot.results.un, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- erp.coef <- extrMWLCoef(act.results.ps, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- ero.coef <- extrMWLCoef(act.results.os, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- erm.coef <- extrMWLCoef(act.results.mu, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- eru.coef <- extrMWLCoef(act.results.un, -->
<!--                         est.points, -->
<!--                         x.names) -->
<!-- all.mwl.coefs <- list(pra.coef, era.coef, prp.coef, pro.coef, prm.coef, pru.coef, -->
<!--                   erp.coef, ero.coef, erm.coef, eru.coef) -->
<!-- names(all.mwl.coefs) <- c('pra', 'era', 'prp', 'pro', 'prm', 'pru',  -->
<!--                           'erp', 'ero', 'erm', 'eru') -->

<!-- ``` -->

<!-- Finally, these coefficients are converted to a surface.  -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!-- makeMWLSurfs <- function(mwl.coefs, -->
<!--                          est.points){ -->

<!--   # Make capture list -->
<!--   surf.list <- list() -->

<!--   mwl.coefs <- mwl.coefs[ ,!names(mwl.coefs) %in% c('lat', 'long')] -->

<!--  # Create surfaces -->
<!--   for(sl in 1:ncol(mwl.coefs)){ -->

<!--    # If NAs -->
<!--    to.cut <- which(is.na(mwl.coefs[,sl])) -->
<!--    if(length(to.cut) > 0){ -->
<!--      surf.points <- est.points[-to.cut,] -->
<!--      surf.value <- mwl.coefs[-to.cut, sl] -->
<!--    } else { -->
<!--      surf.points <- est.points -->
<!--      surf.value <- mwl.coefs[ , sl] -->
<!--    } -->

<!--    # Surface estimation -->
<!--    surf.list[[sl]] <- point2Surface(surf.points,  -->
<!--                                     surf.value, -->
<!--                                     .01,  -->
<!--                                     1.5) -->
<!--   } -->

<!--   names(surf.list) <- names(mwl.coefs) -->

<!-- return(surf.list)     -->
<!-- } -->

<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!-- pra.surfs <- makeMWLSurfs(pra.coef, -->
<!--                           est.points) -->
<!-- era.surfs <- makeMWLSurfs(era.coef, -->
<!--                           est.points) -->

<!-- prp.surfs <- makeMWLSurfs(prp.coef, -->
<!--                           est.points) -->
<!-- pro.surfs <- makeMWLSurfs(pro.coef, -->
<!--                           est.points) -->
<!-- prm.surfs <- makeMWLSurfs(prm.coef, -->
<!--                           est.points) -->
<!-- pru.surfs <- makeMWLSurfs(pru.coef, -->
<!--                           est.points) -->

<!-- erp.surfs <- makeMWLSurfs(erp.coef, -->
<!--                           est.points) -->
<!-- ero.surfs <- makeMWLSurfs(ero.coef, -->
<!--                           est.points) -->
<!-- erm.surfs <- makeMWLSurfs(erm.coef, -->
<!--                           est.points) -->
<!-- eru.surfs <- makeMWLSurfs(eru.coef, -->
<!--                           est.points) -->
<!-- all.mwl.surfs <- list(pra.surfs, era.surfs, prp.surfs, pro.surfs, prm.surfs, pru.surfs, -->
<!--                       erp.surfs, ero.surfs, erm.surfs, eru.surfs) -->
<!-- names(all.mwl.surfs) <- c('pra', 'era', 'prp', 'pro', 'prm', 'pru', -->
<!--                           'erp', 'ero', 'erm', 'eru') -->
<!-- ``` -->