### 2018-05-26 scripts/1_str_preprocessing.Rmd
Update: Str Preprocessing Script to include relative paths.

---
title: "Pre Processing of the Short Term Data"
subtitle:  "For the 'To Airbnb' paper"
author: "Andy Krause and Gideon Aschwanden"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

This script executes a number of very long running data preparation process necessary to analyze the short term rental (Airbnb) data.  If you re-knit this file the code will not evaluate.  To actually evaluate the code you must manually re-run the code chunks individually in RStudio.

```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE, eval=FALSE)

```

## Preliminary Commands

We begin by setting stringsAsFactors to FALSE to avoid unnecessary factor conversions

```{r saf}

 options(stringsAsFactors=FALSE)

```

Next, we load the necessary libraries

```{r load_libraries, warnings=FALSE, message=FALSE}

  library(plyr)
  library(stringr)
  library(lubridate)
  library(dplyr)
  library(multidplyr) # Note you have to instal from the github site hadley/multidplyr
  library(tidyr)
  library(tibble)
  library(magrittr)

```

We then set the paths to the data

```{r set_paths, echo=TRUE}

  source(file.path(getwd(), 'functions', "abb_Functions.R"))

```

Finally, we set the date of analysis (Sept 1, 2015) to help limit the data size

```{r anl_date}

  anl_date <- as.Date('2015-09-01')

```

## Load and Combine Data

Next, we load the data. There are two files for each the property level data as well as the daily observation data.

```{r load_data}

  # Property information
  str_prop_1 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_prop_1.csv'), header=T)
  str_prop_2 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_prop_2.csv'), header=T)

  # Daily rental information
  str_daily_1 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_daily_1.csv'), header=T)
  str_daily_2 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_daily_2.csv'), header=T)

```

We then combine the two files for each type, standardize the field names and select only those fields which we need going forward.  We add a data origin field to the property level information so that we can take the most recent observation in later de-duplication processes.

```{r tcr_data}

  # Property data
  str_df <- dplyr::bind_rows(str_prop_1 %>% 
                               dplyr::mutate(origin = 1), 
                             str_prop_2 %>% 
                               dplyr::mutate(origin = 2)) %>%
    dplyr::select(property_id=Property.ID, host_id=Host.ID, listing_title=Listing.Title, 
                  property_type=Property.Type, listing_type=Listing.Type,
                  created_date=Created.Date, latitude=Latitude, longitude=Longitude,
                  rating=Overall.Rating, max_guests=Max.Guests, 
                  canc_policy=Cancellation.Policy, cleaning_fee=Cleaning.Fee, 
                  nightly_rate=Published.Nightly.Rate, 
                  weekly_rate=Published.Weekly.Rate, monthly_rate=Published.Monthly.Rate, 
                  avg_nightly_rate=Average.Daily.Rate, min_stay=Minimum.Stay, 
                  bedrooms=Bedrooms, bathrooms=Bathrooms, origin)

  # Clean up
  rm(str_prop_1); rm(str_prop_2)

  # Daily data
  daily_df <- dplyr::bind_rows(str_daily_1, str_daily_2) %>%
    dplyr::select(property_id=Property.ID, date=Date, status=Status, price=Price,
                  book_date=Booked.Date, resv_id=Reservation.ID)

  # Clean up
  rm(str_daily_1); rm(str_daily_2);gc()

```

## Prepare Data

Due to our combining of data, we have many duplicate observations.  We now remove any duplicates.  Note that the daily data in its raw format does not have a unique identifier, so we much first create one by combining the property id with the date of the observation.

```{r uniq_ids}

  # Property data
  str_df <- str_df %>% 
    dplyr::arrange(desc(origin)) %>%
    dplyr::filter(!duplicated(property_id)) %>%
    dplyr::select(-origin)

  # Daily data
  daily_df <- daily_df %>%
    dplyr::mutate(uniq_id = paste0(property_id, '_', date)) %>%
    dplyr::filter(!duplicated(uniq_id))

```

The date fields read in as characters, so we also must convert these to an R date format.

```{r fix_dates}

  # Daily data
  daily_df$date <- lubridate::as_date(daily_df$date)
  daily_df$book_date <- lubridate::as_date(daily_df$book_date)

  # Property level data
  str_df$created_date <- lubridate::as_date(str_df$created_date)
  str_df$created_year <- lubridate::year(str_df$created_date)

```

Next, we trim the daily data by date, limiting the observations to the one month period from September 1 2015 to August 31, 2016. We'll eliminate the property-level observations which are outside of the range through a future join. 
```{r date_trim}

  daily_df <- daily_df %>%
    dplyr::filter(date >= anl_date &
                    date <= anl_date + 365)

```

For each property in the short term data, we then calculate some summary statistics of its listing history such as the earliers and latest date for each, the total days on Airbnb, number of blocked periods, its occupancy/available/blocked rate, median length of block period and total number of bookings.

To speed up this calculate, we compute these summary statistics in parallel.  Total run time will depend on your system.

```{r summ_stats}

  # Create new cluster
  abb_cl <- get_default_cluster()

  # Register function and variable
  cluster_copy(abb_cl, abbCalcBookStr)

  # Summarize daily by property
  daily_summ <- daily_df %>% 
    multidplyr::partition(property_id, cluster=abb_cl) %>% 
    dplyr::do(abbCalcBookStr(.)) %>% 
    dplyr::collect()

```

We then join the summary stats to the property-level data with an inner join, meaning that any properties without oservations in this period are removed.

```{r join_summ}

  str_df <- str_df %>%
    dplyr::inner_join(daily_summ, 
                      by='property_id')

```

In order to store all of our data efficiency in a single object, we now **nest** the daily observations for each property into a list-column (each property observation will contain a data.frame of daily observations).

```{r join_nest}

  # Nest daily data
  daily_tdf <- daily_df %>%
    as.tibble() %>%
    multidplyr::partition(property_id, cluster=abb_cl) %>% 
    dplyr::do(tidyr::nest(., -property_id, .key='daily_data')) %>% 
    dplyr::collect()

  # Add nested daily data to structural data
  str_df <- str_df %>%
    as.tibble() %>%
    left_join(daily_tdf, by='property_id')

```

### Export Data

Finally, we write out the date to an RDS object for future use.

```{r write_out}

 dir.create(file.path(getwd(), 'data', 'prepared'))
 saveRDS(str_df, file=file.path(getwd(), 'data', 'prepared', 'str_data.RDS'))

```