### 2018-05-30 scripts/5_data_viz.Rmd
Image updates

---
title: "5_data_viz"
author: "Andy Krause"
date: "May 10, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE)

```

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tmap)
  library(magrittr)
  library(dplyr)
  library(ggplot2)

```

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, results='hide'}

  load(file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))

  dir.create(file.path(getwd(), 'images'))

  str_col = col_df$rgb[1]

```

## Suburb Location Map

```{r sub_loc.map, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  # Set submarket colors
  suburb_sf$col <- 'gray80'
  suburb_sf$col[suburb_sf$class == 'Middle'] <- 'gray50'
  suburb_sf$col[suburb_sf$class == 'Inner'] <- 'gray20'

   # Make map

  png(filename = file.path(getwd(), 'images', "sub_loc_map.png"), 
       width = 6.5, height = 6.5, units = 'in', res = 300)

  tm_shape(suburb_sf) +
    tm_polygons("class", palette=c('gray20', 'gray50', 'gray80'), title='Suburb Class') +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=5) +
    tm_add_legend(type='fill', border.lwd=3, col='white', border.col='black',
                  title='Study Area')

  dev.off()

```

## Point Location Maps

```{r location map}

  sm_col <- c(col_df$rgb[3], col_df$rgb[4], col_df$rgb[5], 
              col_df$rgb[6], col_df$rgb[1])

  png(filename = file.path(getwd(), 'images', "ltr_loc_map.png"), 
       width = 4, height = 4, units = 'in', res = 300)
  tm_shape(studyarea_sf) + 
    tm_borders(col='gray50', lwd=2) +
  tm_shape(suburb_sf %>% dplyr::filter(in_studyarea == 1)) +
    tm_borders(col='gray90', lwd=.5) +
  tm_shape(ltr_sf %>% dplyr::rename(Submarket = geo_mrkt)) +
    tm_symbols('Submarket', size=.02,
               palette = sm_col, alpha=.3, shape=16, legend.col.show=TRUE)
  dev.off()

  png(filename = file.path(getwd(), 'images', "str_loc_map.png"), 
       width = 4, height = 4, units = 'in', res = 300)
  tm_shape(studyarea_sf) + 
    tm_borders(col='gray50', lwd=2) +
  tm_shape(suburb_sf %>% dplyr::filter(in_studyarea == 1)) +
    tm_borders(col='gray90', lwd=.5) +
  tm_shape(str_sf %>% dplyr::rename(Submarket = geo_mrkt)) +
    tm_symbols('Submarket', size=.02,
               palette = sm_col, alpha=.3, shape=16)
  dev.off()

```

```{r smtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.width=9}

   # Make Airbnb submarket map

   strtype_bar <- ggplot() +
     geom_bar(data=str_tdf, aes(x=geo_mrkt, fill=geo_mrkt)) +
     scale_fill_manual(values=sm_col, name='Submarket') +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'bottom',
           legend.title = element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank()) +
     ggtitle('Short Term Property Type\nBy Submarket')+
     theme(plot.title = element_text(hjust = 0.5))

   # Make Long term rental map

   ltrtype_bar <- ggplot() +
     geom_bar(data=ltr_tdf, aes(x=geo_mrkt, fill=geo_mrkt)) +
     scale_fill_manual(values=sm_col) +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'bottom',
           legend.title = element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank()) +
     ggtitle('Long Term Property Type\nBy Submarket')+
     theme(plot.title = element_text(hjust = 0.5))

     png(filename = file.path(getwd(), 'images', "loc_bar.png"), 
       width = 8, height = 4, units = 'in', res = 300)

     ggMultiPlots(strtype_bar, ltrtype_bar, cols=2)
     dev.off()

```

```{r maps}

   dens_cols <- c(col_df$rgb[3], col_df$rgb[1], col_df$rgb[2])

   # Add together for plotting
   rev_df <- bind_rows(str_tdf %>% 
                         dplyr::select(type, revenue=lik_revenue_net) %>%
                         dplyr::mutate(tenure = 'Short Term (Likely)   ',
                                       alpha = .5),
                       str_tdf %>% 
                         dplyr::select(type, revenue=pot_revenue_net) %>%
                         dplyr::mutate(tenure = 'Short Term (Potential)   ',
                                       alpha = .4),
                       ltr_df %>%
                         dplyr::select(type, revenue=gross_revenue) %>%
                         dplyr::mutate(tenure = 'Long-Term   ',
                                       alpha = .9))

  png(filename = file.path(getwd(), 'images', "rev_dens.png"), 
       width = 5.5, height = 4, units = 'in', res = 300)

   ggplot(rev_df,
          aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=dens_cols) +
    scale_color_manual(values=dens_cols) +
    ggtitle('Short Term Revenue vs Long Term Revenue') +
    xlab('Annual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20,000', '$40,000', '$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000))

    dev.off()

```

Host type investigation

```{r host_types}

  str.col <- col_df$rgb[1]

  ## Create table of host block frequencies

  host.block.freq <- ggplot(str_tdf, aes(x=block_rate)) +
     geom_density(fill=str.col, color=str.col) +
     ylab('Frequency\n') +
     xlab('\n% of Time Property is Blocked') +
     scale_x_continuous(breaks=seq(0,1,.2),
                        labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
     #ggtitle('Host Block Rate Frequency\n') +
     theme(plot.title = element_text(hjust = 0.5)) +
     theme(axis.text.y=element_blank(),
           axis.ticks.y=element_blank())

  block.rate.period <- ggplot(str_tdf, aes(x=block_rate, y=nbr_block * (366/total_days))) +
    geom_point(fill=str.col, color=str.col, size=.2) +
    ylab('Nbr of Blocked Periods\n') +
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    #ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

    png(filename = file.path(getwd(), 'images', "block_rate.png"), 
       width = 5.5, height = 3, units = 'in', res = 300)

       ggMultiPlots(host.block.freq, block.rate.period, cols=2)

    dev.off()

```

```{r host_types2}

  base.host <- ggplot(str_tdf,
                    aes(x=block_rate, y=blockpertime)) +
                geom_point(alpha=.7, size=.4, shape=16, color='gray70') +
                ylab('# Blckd Periods\n') +
                xlab('\n% of Time Property is Blocked') +
                ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
                scale_x_continuous(breaks=seq(0, 1, .2),
                                   labels=c('0%', '20%', '40%',
                                            '60%', '80%', '100%')) +
                #scale_color_manual(values=c(abb.col[2], abb.col[4],
                #                            abb.col[7], abb.col[9])) +
                theme(legend.position='bottom',
                      legend.title = element_blank(),
                      plot.title = element_text(hjust = 0.5))+
                guides(colour = guide_legend(override.aes = list(size=3,                                                                       alpha=1)))

   ps.host <- base.host +
    geom_point(data=str_tdf[str_tdf$host_type == 'Profit Seeker', ], size=.4)  +
    ggtitle('Profit Seekers')

  os.host <- base.host +
    geom_point(data=str_tdf[str_tdf$host_type == 'Opportunistic Sharer', ],
               size=.4)  +
    ggtitle('Opportunistic Sharer')

  mu.host <- base.host +
    geom_point(data=str_tdf[str_tdf$host_type == 'Multi-Platform User', ],
               size=.4)  +
    ggtitle('Multi-Platform User')

  un.host <- base.host +
    geom_point(data=str_tdf[str_tdf$host_type == 'Unknown', ], size=.4)  +
    ggtitle('Unknown')

  png(filename = file.path(getwd(), 'images', "host_scat.png"), 
       width = 5.5, height = 3, units = 'in', res = 300)

    ggMultiPlots(ps.host, mu.host, os.host, un.host, cols=2)

  dev.off()

```

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=4, fig.cap='\\label{htc}Host Type Counts'}

 ## Build plot

  png(filename = file.path(getwd(), 'images', "host_bar.png"), 
       width = 3, height = 3, units = 'in', res = 300)

 ggplot(str_tdf, aes(host_type, fill=host_type)) +
    geom_bar() +
    scale_fill_manual(values=c(col_df$rgb[2], col_df$rgb[4], col_df$rgb[7], 
                               col_df$rgb[9]))+
    theme(legend.position='none') +
    ylab('Frequency\n') +
    xlab('\nHost Type') +
    scale_x_discrete(labels=c('Profit','Sharer','Multi','Unknown'))+
    #ggtitle('short-term Host Types') +
    theme(plot.title = element_text(hjust = 0.5))

  dev.off()

```