### 2018-11-04 scripts/4_data_analysis.Rmd

---
title: "Data Analysis and Modeling"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

### Introduction

This document describes all of the code used to the generate the output included in the 'To Airbnb?' study. This analysis uses the output from the **3_data_prep.rmd** file found in the same online repository.

## Preliminary Commands

&nbsp;

We begin with a number of preliminary commands.  First we load all of the necessary R libraries for this analysis.

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)

```

```{r set_options, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

```

Next, we source the necessary custom code files for the analysis.

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  source(file.path(getwd(), 'functions', 'abb_Functions.R'))

```

### Set Parameters

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  # Exchange rate: AUS to US 
  exch_rate <- 1.32

  # Date of investment decision
  decision_date <- lubridate::as_date('2015-09-01')

```

## Load Data

We then load the data.

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  load(file.path(getwd(), 'data', 'prepared', 'prepared_data.RData'))
  str_col = col_df$rgb[1]

```

#### Compare Revenue Distributions

```{r flatten_ltr}

  ltr_df <- ltr_tdf %>% 
    unnest() %>%
    select(-suburb.x) %>%
    rename(suburb=suburb.y)

```

Before we examine the distributions of the observed short-term revenues and the estimated long-term revenues, we must produce the estimates of long-term revenue for the short-term properties.

We start by specifying the imputation model specification we will use to estimated likely long-term (weekly) rental prices for each short-term property.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr_mod_spec <- formula(log(price) ~ as.factor(type) +
                            as.factor(bedbath) + as.factor(suburb) +
                            as.factor(month))

```

We then employ a custom function to impute the long term rents for the short-term properties.  The 'clip.field' is a spatial fixed effects that add spatial recognition to the models.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$tbg <- paste0(str_tdf$product,'_', str_tdf$geo_mrkt)
  ltr_df$tbg <- paste0(ltr_df$product,'_', ltr_df$geo_mrkt)

  imp_df <- abbImputeCompare(str.df=str_tdf, 
                                ltr.df=ltr_df, 
                                mod.spec=ltr_mod_spec, 
                                match.factor='suburb',
                                split.field='tbg', 
                                verbose=FALSE)

  kimp_df <- knnImputeWrap(ltr_df = ltr_df, 
                           str_df = str_tdf,
                           knns = c(7,7))

  imp_df$ltr_imp_revenueK <- kimp_df$year[match(imp_df$property_id,
                                               kimp_df$property_id)]
  imp_df <- imp_df %>%
    rename(ltr_imp_revenueH = ltr_imp_revenue) %>%
    mutate(ltr_imp_revenueK = ((365-imp_dom)/365) * ltr_imp_revenueK,
           ltr_imp_revenue = .5 * (ltr_imp_revenueH + ltr_imp_revenueK))
  imp_df <- imp_df %>%
    left_join(str_tdf %>%
                dplyr::select(property_id, bedrooms, bathrooms, type, latitude,
                              longitude),
              by='property_id')

    save(imp_df, file=file.path(getwd(), 'data', 'analyzed', 'reviewers.RData'))

```

From the imputed results, we extract the imputed rental values and merge to the original short-term data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df %>%
                       dplyr::select(-c(bedrooms, bathrooms, type, latitude, longitude)),
                     by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(revenue_net > ltr_imp_revenue, 1, 0),
                  # lik_pref = ifelse(lik_revenue_net > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue_net > ltr_imp_revenue, 1, 0),
                  pot_pref_mc = ifelse(pot_revenue_net * .85 > ltr_imp_revenue, 1, 0),
                  pot_pref_us = ifelse(pot_revenue_net > ltr_imp_revenue * .85, 1, 0))

```

## Logit models

Now we attempt to explain the drivers of preference with a set of logistic regression models.

First, a basic model with just structural characteristics.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_pot_str <- glm(pot_pref~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
  glm_pot_str_mc <- glm(pot_pref_mc~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
  glm_pot_str_us <- glm(pot_pref_us~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)

```

Then a model that adds in the geographic submarkets

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_pot_loc <- glm(pot_pref ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
  glm_pot_loc_mc <- glm(pot_pref_mc ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
  glm_pot_loc_us <- glm(pot_pref_us ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
```

Finally a model that adds in host policies.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_pot_host <- glm(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + I(canc_policy == 'Flexible'),
                      family=binomial(link='logit'),
                      data=str_tdf)
  glm_pot_host_mc <- glm(pot_pref_mc ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + I(canc_policy == 'Flexible'),
                      family=binomial(link='logit'),
                      data=str_tdf)
  glm_pot_host_us <- glm(pot_pref_us ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + I(canc_policy == 'Flexible'),
                      family=binomial(link='logit'),
                      data=str_tdf)

  pot_glms <- list(str=glm_pot_str,
                   loc=glm_pot_loc,
                   host=glm_pot_host)
  pot_mc_glms <- list(str=glm_pot_str_mc,
                   loc=glm_pot_loc_mc,
                   host=glm_pot_host_mc)
  pot_us_glms <- list(str=glm_pot_str_us,
                   loc=glm_pot_loc_us,
                   host=glm_pot_host_us)

```

<!-- ### Host types -->

<!-- Finally, we test to see if grouping by hosts types has any impact.   -->

<!-- First we must divide hosts up.  We start by looking at block rate.   -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   ggplot(str_tdf, aes(x=block_rate)) +  -->
<!--     geom_density(fill=str_col, color=str_col) + -->
<!--     ylab('Frequency\n') +  -->
<!--     xlab('\n% of Time Property is Blocked') + -->
<!--     scale_x_continuous(breaks=seq(0,1,.2), -->
<!--                        labels=c('0%', '20%', '40%', '60%', '80%', '100%')) + -->
<!--       ggtitle('Host Block Rate Frequency\n') + -->
<!--       theme(plot.title = element_text(hjust = 0.5)) + -->
<!--       theme(axis.text.y=element_blank(), -->
<!--           axis.ticks.y=element_blank()) -->

<!-- ``` -->

<!-- We then add the number of blocked periods -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   ggplot(str_tdf, aes(x=block_rate, y=nbr_block * (366/total_days))) +  -->
<!--     geom_point(fill=str_col, color=str_col, size=.2) + -->
<!--     ylab('Nbr of Blocked Periods\n') +  -->
<!--     xlab('\n% of Time Property is Blocked') + -->
<!--     scale_x_continuous(breaks=seq(0, 1, .2), -->
<!--                        labels=c('0%', '20%', '40%', '60%', '80%', '100%')) + -->
<!--     ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') + -->
<!--     theme(plot.title = element_text(hjust = 0.5))  -->

<!-- ``` -->

<!-- No real pattern emerges, so we divide manually.  -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   str_tdf$blockpertime <- round(str_tdf$nbr_block * (366 / str_tdf$total_days), 1) -->

<!--   str_tdf$host_type <- 'Unknown' -->
<!--   str_tdf$host_type[str_tdf$blockpertime > 12] <- 'Multi-Platform User' -->
<!--   str_tdf$host_type[str_tdf$block_rate <= .25] <- 'Profit Seeker' -->
<!--   str_tdf$host_type[str_tdf$block_rate >= .75] <- 'Opportunistic Sharer' -->
<!--   str_tdf$host_type <- factor(str_tdf$host_type, -->
<!--                               levels=c('Profit Seeker', 'Opportunistic Sharer',  -->
<!--                                         'Multi-Platform User', 'Unknown')) -->

<!--  ggplot(str_tdf,  -->
<!--        aes(x=block_rate, y=blockpertime, color=host_type)) +  -->
<!--   geom_point(alpha=.5, size=.8, shape=16) + -->
<!--   ylab('Number of Blocked Periods\n') +  -->
<!--   xlab('\n% of Time Property is Blocked') + -->
<!--       ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') + -->

<!--     scale_x_continuous(breaks=seq(0, 1, .2), -->
<!--                      labels=c('0%', '20%', '40%', '60%', '80%', '100%')) + -->
<!--     scale_color_manual(values=c(col_df$rgb[2], col_df$rgb[4], col_df$rgb[7],  -->
<!--                                 col_df$rgb[9])) +  -->
<!--     theme(legend.position='bottom', -->
<!--            legend.title = element_blank(), -->
<!--          plot.title = element_text(hjust = 0.5))+ -->
<!--      guides(colour = guide_legend(override.aes = list(size=3, -->
<!--                                                      alpha=1))) -->

<!-- ``` -->

<!-- And re-run the logistic regression models.  -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   ht_. <- split(str_tdf, str_tdf$host_type) -->

<!--   lik_spec <- as.formula(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) + -->
<!--                           guest_per_bed + min_stay + I(canc_policy == 'Flexible')) -->

<!--   pot_spec <- as.formula(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) + -->
<!--                           guest_per_bed + min_stay + I(canc_policy == 'Flexible')) -->

<!--   lik_ht_glms <- purrr::map(ht_., -->
<!--                          .f=glm, -->
<!--                          formula=lik_spec, -->
<!--                          family=binomial(link='logit')) -->

<!--   pot_ht_glms <- purrr::map(ht_., -->
<!--                          .f=glm, -->
<!--                          formula=pot_spec, -->
<!--                          family=binomial(link='logit')) -->

<!-- ``` -->

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  dir.create(file.path(getwd(), 'data', 'analyzed'))
  save(str_tdf, ltr_tdf, ltr_df, col_df, suburb_sf, studyarea_sf, sa1_sf, str_sf, ltr_sf,
       pot_glms, core_suburbs, str_tdf_all, pot_mc_glms, pot_us_glms,
       #pot_ht_glms, lik_ht_glms,
       file=file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))

  save(ltr_df, file=file.path(getwd(), 'data', 'analyzed', 'reviewers.RData'))

```