### 2018-11-10 scripts/5_data_viz.Rmd
Fix beach data issue, join maps, simplify data preparation discussion

---
title: "Data Visualization"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE)

```

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tmap)
  library(magrittr)
  library(dplyr)
  library(ggplot2)

```

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, results='hide'}
  load(file.path(getwd(), 'data', 'prepared', 'prepared_data.RData'))

  load(file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))

  dir.create(file.path(getwd(), 'images'))

  str_col = col_df$rgb[1]

```

## Suburb Location Map

```{r sub_loc.map, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  # Set submarket colors
  suburb_sf$col <- 'gray80'
  suburb_sf$col[suburb_sf$class == 'Middle'] <- 'gray50'
  suburb_sf$col[suburb_sf$class == 'Inner'] <- 'gray20'

  inn_sf <- suburb_sf %>% dplyr::filter(class == 'Inner') %>% st_union()
  mid_sf <- suburb_sf %>% dplyr::filter(class == 'Middle' & 
                                          name != 'Seabrook' &  
                                           name != 'Altona Meadows') %>% st_union()

  sub_sdf <- ltr_tdf %>%
    group_by(suburb.x, Submarket) %>%
    dplyr::summarize(count=n())
  sub_sf <- suburb_sf %>%
    dplyr::left_join(sub_sdf %>%
                       dplyr::select(name = suburb.x, Submarket))

  cc_sf <- sub_sf %>% dplyr::filter(Submarket == '   City-Core   ') %>% st_union()
  beach_sf <- sub_sf %>% dplyr::filter(Submarket == '   Beach   ') %>% st_union()

   # Make map

  png(filename = file.path(getwd(), 'images', "sub_loc_map.png"), 
       width = 4.5, height = 4.5, units = 'in', res = 300)

  tm_shape(suburb_sf %>% dplyr::filter(class != 'Outer')) +
    tm_polygons("class", palette=c('gray50', 'gray80'), title='Suburb Class') +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=5) +
    tm_add_legend(type='fill', border.lwd=3, col='white', border.col='black',
                  title='Study Area')+
  tm_layout(
    legend.position = c("left","bottom"),
    legend.text.size = 0.5)

  dev.off()

```

## Point Location Maps

```{r location map}

  sm_col <- c(col_df$rgb[3], col_df$rgb[4], col_df$rgb[5], 
              col_df$rgb[6])

  ltr_map <-   tm_shape(mid_sf) + 
    tm_fill(col='gray95', lwd=1) +
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(inn_sf) + 
    tm_fill(col='gray85', lwd=1) +  
  tm_shape(cc_sf) + 
    tm_fill(col='gray75', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_fill(col='gray65', lwd=1) +
  tm_shape(suburb_sf %>% dplyr::filter(in_studyarea == 1)) +
    tm_borders(col='gray70', lwd=.5) +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=2) +
  tm_shape(ltr_sf %>% dplyr::rename(Submarket = geo_mrkt)) +
    tm_symbols('Submarket', size=.02,
               palette = sm_col, alpha=.9, shape=16, legend.col.show=TRUE)+
    tm_layout(title = "LTR Observations",
              title.position = c("left", "top"),
              title.size = .8) +
  tm_shape(inn_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(cc_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_borders(col = 'gray40', lwd=1)

  str_map <-   tm_shape(mid_sf) + 
    tm_fill(col='gray95', lwd=1) +
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(inn_sf) + 
    tm_fill(col='gray85', lwd=1) +  
  tm_shape(cc_sf) + 
    tm_fill(col='gray75', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_fill(col='gray65', lwd=1) +
  tm_shape(suburb_sf %>% dplyr::filter(in_studyarea == 1)) +
    tm_borders(col='gray70', lwd=.5) +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=2) +
  tm_shape(str_sf %>% dplyr::rename(Submarket = geo_mrkt)) +
    tm_symbols('Submarket', size=.02,
               palette = sm_col, alpha=.9, shape=16, legend.col.show=TRUE)+
    tm_layout(title = "LTR Observations",
              title.position = c("left", "top"),
              title.size = .8) +
  tm_shape(inn_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(cc_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_borders(col = 'gray40', lwd=1)

  sltr_map <- tmap_arrange(ltr_map, str_map)

  png(filename = file.path(getwd(), 'images', "ltr_loc_map.png"), 
       width = 7, height = 3.5, units = 'in', res = 300)
    sltr_map
  dev.off()

```

```{r smtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.width=9}

   # Make Airbnb submarket map
   str_tdf$Submarket <- as.character(str_tdf$geo_mrkt)
   str_tdf$Submarket[str_tdf$Submarket == 'city-core'] <- '   City-Core   '
   str_tdf$Submarket[str_tdf$Submarket == 'city'] <- '   City   '
   str_tdf$Submarket[str_tdf$Submarket == 'beach'] <- '   Beach   '
   str_tdf$Submarket[str_tdf$Submarket == 'suburban'] <- '   Suburban   '
   str_tdf$Submarket <- factor(str_tdf$Submarket, 
                                  levels = c('   City-Core   ', '   City   ', 
                                             '   Suburban   ', '   Beach   '))

   strtype_bar <- ggplot() +
     geom_bar(data=str_tdf, aes(x=Submarket, fill=Submarket)) +
     scale_fill_manual(values=sm_col, name='Submarket') +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'none',
           legend.title = element_blank(),
           axis.ticks.x=element_blank()) +
     ggtitle('Short Term Properties')+
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=.5)) + 
     labs(caption = paste0('\nBreakdown of 2,415 short-term properties by submarket and ',
                           ' structure type.\n'))
  # strtype_bar

   # Make Long term rental map

   ltr_tdf$Submarket <- as.character(ltr_tdf$geo_mrkt)
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'city-core'] <- '   City-Core   '
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'city'] <- '   City   '
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'beach'] <- '   Beach   '
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'suburban'] <- '   Suburban   '
   ltr_tdf$Submarket <- factor(ltr_tdf$Submarket, 
                                  levels = c('   City-Core   ', '   City   ', 
                                             '   Suburban   ', '   Beach   '))
   ltrtype_bar <- ggplot() +
     geom_bar(data=ltr_tdf, aes(x=Submarket, fill=Submarket)) +
     scale_fill_manual(values=sm_col) +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'none',
           legend.title = element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank()) +
     ggtitle('Long Term Properties')+
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=.5)) + 
     labs(caption = paste0('\nBreakdown of 64,305 long-term properties by submarket and ',
                           ' structure type\n'))
     #ltrtype_bar

     png(filename = file.path(getwd(), 'images', "loc_bar.png"), 
       width = 8, height = 4, units = 'in', res = 300)

     ggMultiPlots(strtype_bar, ltrtype_bar, cols=2)
     dev.off()

```

```{r maps}

   dens_cols <- c(col_df$rgb[3], col_df$rgb[1], col_df$rgb[2])

   # Add together for plotting
   rev_df <- bind_rows(str_tdf %>% 
                         dplyr::select(type, revenue=revenue_net) %>%
                         dplyr::mutate(tenure = '  Short Term (Actual)   ',
                                       alpha = .5),
                       str_tdf %>% 
                         dplyr::select(type, revenue=pot_revenue_net) %>%
                         dplyr::mutate(tenure = '  Short Term (Potential)   ',
                                       alpha = .4),
                       ltr_df %>%
                         dplyr::select(type, revenue=gross_revenue) %>%
                         dplyr::mutate(tenure = '  Long-Term   ',
                                       alpha = .9))

  png(filename = file.path(getwd(), 'images', "rev_dens.png"), 
       width = 5.5, height = 4, units = 'in', res = 300)

   ggplot(rev_df,
          aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=dens_cols) +
    scale_color_manual(values=dens_cols) +
    # ggtitle('Revenue Comparison: Short Term vs Long Term Properties') +
    xlab('Annual Revenue (A$)') +
    ylab('')+
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('A$0', 'A$20,000', 'A$40,000', 'A$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank())+
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nComparison of the distribution of observed long-term rental',
                           ' revenues of 64,305 properties (blue) to observed revenues of ', 
                           ' 2,415 short\nterm properties (pink).  The purple curve ',
                           ' represents the potential revenues of the 2,415 short term ',
                           ' properties (includes imputed bookings).')) +
    coord_cartesian(xlim=c(0, 70000))

    dev.off()

```

Host type investigation

```{r host_types}

  str.col <- col_df$rgb[1]

  ## Create table of host block frequencies

  host.block.freq <- ggplot(str_tdf_all, aes(x=block_rate)) +
     geom_density(fill=str.col, color=str.col) +
     ylab('Frequency\n') +
     xlab('\n\n% of Time Property is Blocked') +
     scale_x_continuous(breaks=seq(0,1,.2),
                        labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
     #ggtitle('Host Block Rate Frequency\n') +
     theme(plot.title = element_text(hjust = 0.5, size = 10)) +
     theme(axis.text.y=element_blank(),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nDensity plot of frequency of blocking schedule for short term ',
                           '\nleasing observations. Blocking indicates that the property ',
                           'is \nnot available for nightly rental on Airbnb. Vertical lines',
                           '\ndelineate the three host type groupings.'),
          title = 'Host Blocking Frequency\n') +
    geom_vline(xintercept = .25, linetype="dashed", color = "black", size=.5, alpha=.5)+
    geom_vline(xintercept = .75, linetype="dashed", color = "black", size=.5, alpha=.5)

  block.rate.period <- ggplot(str_tdf_all, 
                              aes(x=block_rate, y=nbr_block * (366/total_days))) +
    geom_point(fill=alpha(str.col,.3), color=alpha(str.col,.3), size=.1, shape=16) +
    ylab('Nbr of Blocked Periods\n') +
    xlab('\n\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    #ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nCumulative time of blocking for short term properties ',
                           '(x-axis) \nversus the total number of periods where the ',
                           'property was \nblocked from nightly rental.',
                           'Vertical lines delineate the three \nhost type groupings.'),
          title = 'Block Rate vs Blocked Periods\n')+
    geom_vline(xintercept = .25, linetype="dashed", color = "black", size=.5, alpha=.5)+
    geom_vline(xintercept = .75, linetype="dashed", color = "black", size=.5, alpha=.5)

    png(filename = file.path(getwd(), 'images', "block_rate.png"), 
       width = 5.5, height = 3, units = 'in', res = 300)

       ggMultiPlots(host.block.freq, block.rate.period, cols=2)

    dev.off()

```

```{r host_types2}

  base.host <- ggplot(str_tdf_all,
                    aes(x=block_rate, y=blockpertime)) +
                geom_point(alpha=.7, size=.4, shape=16, color='gray70') +
                ylab('# Blckd Periods\n') +
                xlab('\n% of Time Property is Blocked') +
                ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
                scale_x_continuous(breaks=seq(0, 1, .2),
                                   labels=c('0%', '20%', '40%',
                                            '60%', '80%', '100%')) +
                #scale_color_manual(values=c(abb.col[2], abb.col[4],
                #                            abb.col[7], abb.col[9])) +
                theme(legend.position='bottom',
                      legend.title = element_blank(),
                      plot.title = element_text(hjust = 0.5))+
                guides(colour = guide_legend(override.aes = list(size=3,                                                                       alpha=1)))

   ps.host <- base.host +
    geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Profit Seeker', ], size=.4)  +
    ggtitle('Profit Seekers')

  os.host <- base.host +
    geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Opportunistic Sharer', ],
               size=.4)  +
    ggtitle('Opportunistic Sharer')

  # mu.host <- base.host +
  #   geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Multi-Platform User', ],
  #              size=.4)  +
  #   ggtitle('Multi-Platform User')

  un.host <- base.host +
    geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Unknown', ], size=.4)  +
    ggtitle('Unknown')

  png(filename = file.path(getwd(), 'images', "host_scat.png"), 
       width = 4, height = 6.5, units = 'in', res = 300)

    ggMultiPlots(ps.host, un.host, os.host, cols=1)

  dev.off()

```

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=4, fig.cap='\\label{htc}Host Type Counts'}

 ## Build plot

  png(filename = file.path(getwd(), 'images', "host_bar.png"), 
       width = 3, height = 3, units = 'in', res = 300)

 ggplot(str_tdf_all, aes(host_type, fill=host_type)) +
    geom_bar() +
    scale_fill_manual(values=c(col_df$rgb[2], col_df$rgb[4], col_df$rgb[9]))+
    theme(legend.position='none') +
    ylab('') +
    xlab('') +
    scale_x_discrete(labels=c('Profit-Seekers','Opportunistic Sharers','Unknowns'))+
    # ggtitle('short-term Host Types') +
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nBreakdown of Host Types for Short Term Properties:\n\n',
                           '1) Profit Seekers:               <= 25% Blocked             3,056 Obs.\n',
                           '2) Opportunistic Sharers:   >= 75% Blocked             1,743 Obs.\n',
                           '3) Unknowns:                     25% > 75% Blocked       2,707 Obs.\n')) + 
   annotate("text", x = 1, y = 3300, label = "41%") +
   annotate("text", x = 2, y = 2000, label = "23%") +
   annotate("text", x = 3, y = 2949, label = "36%") +
   coord_cartesian(ylim=c(0,3500))

  dev.off()

```

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

  library(ggplot2)

   load(file.path('c:/code', 'research', 'airbnbMelbourne',
                  'data', 'analyzed', 'reviewers.RData'))

  imp_df$bed_lab <- paste0(imp_df$bedrooms, ' Beds')
  fig_cap <- paste0('This figure compares the estimated long term rent for short term rental properties via the hedonic imputation\nmethod (X-axis) and the matched pairs-type method (Y-axis).  It is broken down by bedroom count and\nstructure type (Apartment vs House). Overall correlation is 89.4%')

  # imp_cor <- imp_df %>%
  #   dplyr::group_by(type, bed_lab) %>%
  #   dplyr::summarize(count = n(),
  #                    cor = cor(ltr_imp_revenueH, ltr_imp_revenueK))
  # 
  #   
  # d1 + geom_text(data=data.frame(x=0.25, y=1.5e+04, label=1:7, 
  #                              color=c("D","E","F","G","H","I","J")), 
  #              aes(x,y,label=label), inherit.aes=FALSE)

  png(filename = file.path(getwd(), 'images', "imp_cor.png"), 
       width = 7, height = 4, units = 'in', res = 300)

  ggplot(imp_df,
         aes(x=ltr_imp_revenueH, y=ltr_imp_revenueK, col = type)) + 
    geom_point(alpha = .5, size=.5) + 
    facet_grid(type ~ bed_lab) + 
    xlab('\nHedonic Revenue (A$)\n') + 
    ylab('Matched Pairs Revenue (A$)\n') + 
    theme(legend.position = 'none',
          plot.caption=element_text(hjust = 0),
          plot.title = element_text(hjust = 0.5)) +
    scale_color_manual(values = c(col_df$rgb[8], col_df$rgb[10])) + 
    scale_x_continuous(breaks = c(15000, 30000, 45000),
                       labels = c('$15k', '$30k', '$45k'))+
    scale_y_continuous(breaks = c(15000, 30000, 45000),
                       labels = c('$15k', '$30k', '$45k'))+
    geom_abline(intercept = 0, slope = 1, linetype = 2,
                color = 'gray50') +
    labs(caption = fig_cap)

  dev.off()

```