### 2018-11-23 papers/round_3_reviews.Rmd

---
title: "To Airbnb?: Factors impacting short-term leasing preference"
subtitle: 'Reviewer Responses'
date: "11/8/2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

## What's Changed

1. Bolstered the literature review
* Provided larger coverage of the short-term rental issues by extending discussion to some non-Real Estate centric work [@kakar2018visible; @cheng2018sharing]
* Added in a few very recent publications to help solidify the importance of this work [@manning2018hotel] and to provide deeper look at previous short term rental pricing work[@gibbs2018use; @gibbs2018pricing; @oskam2018s]

## References