### 2018-11-28 papers/round_3_reviews.Rmd
Revision #3

---
title: "To Airbnb?: Factors impacting short-term leasing preference"
subtitle: 'Review Update'
date: "11/25/2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

## What's Changed

#### 1. Bolstered the literature review (pgs 2-5)

* Provides larger coverage of the short-term rental issues by extending discussion to some non-Real Estate centric work [@kakar2018visible; @cheng2018sharing]

* Adds in a few very recent publications to help solidify the importance of this work [@manning2018hotel] and to provide deeper look at previous short term rental pricing work[@gibbs2018use; @gibbs2018pricing; @oskam2018s]

&nbsp; 
&nbsp;

#### 2. Added Conditional Expectations Section (pgs 24 - 26)

* Provides a clearer look at how the various factor influence short-term preference

* Provides more explicit advise on which situations and strategies are most promising for investors looking for short-term preferable properties.

&nbsp; 
&nbsp;

## References